### 2024-10-16 notebooks/02-crosstables.qmd

---
title: Crosstabulations
author: Isaiah
format: 
  html: 
    code-fold: false
    code-tools: true
    code-overflow: wrap
    code-copy: true
    code-annotations: true
---

## Crosstables

```{r}
#| label: tbl-q19
#| tbl-cap: "Confidence vote counts will be accurate in Maricopa County, AZ by Experiment Condition"

# Crosstab of Q19 by treatment group
df |> 
  janitor::tabyl(group, q19, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q19') |>
  gt::gt(
    rowname_col = 'Q19',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q19') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Q19. Vote Count Confidence for Maricopa County, AZ"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q19. How confident are you that votes in Maricopa County, AZ will be counted as voters intend in the elections this November?",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# df |> 
#   janitor::tabyl(group, q19, show_na = F) |>  
#   janitor::adorn_totals('both', na.rm = T) |> 
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |> 
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'Q19') |> 
#   kableExtra::kbl() |> 
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}
#| label: tbl-q21
#| tbl-cap: "Election Staff Commitment to Fairness and Accuracy by Experiment Condition"
#| tbl-cap-location: top

df |> 
  janitor::tabyl(group, q21, show_na = F) |> 
  # janitor::adorn_totals("both") |> 
  janitor::adorn_percentages("row") |> 
  janitor::adorn_pct_formatting(digits = 2, affix_sign = F) |>
  janitor::adorn_ns("rear") |> 
  janitor::adorn_title("combined", 
                       row_name = "Group",
                       col_name = "Q21") |>
  # bind_rows(.id = "group_var") |>   
  gt::gt(
    rowname_col = "row_var",
    groupname_col = "group_var") |>   
  gt::cols_label(
    c('Group/Q21') ~ "Group") |>
  # gt::cols_label_with(columns = '...1', fn = ~sjlabelled::get_label(df$q5.clps)) |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Q21. Election staff commitment to fairness and accuracy"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q21. Think about the election staff and volunteers who handle the administration and conduct of elections in Maricopa County, AZ. How committed do you think they will be to making sure the elections held this November are fair and accurate?",
    locations = gt::cells_column_spanners()) |>
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# df |> 
#   janitor::tabyl(group, q21, show_na = F) |> 
#   janitor::adorn_totals("both") |> 
#   janitor::adorn_percentages("row") |> 
#   janitor::adorn_pct_formatting(digits = 2, affix_sign = F) |>
#   janitor::adorn_ns("rear") |> 
#   janitor::adorn_title("combined", 
#                        row_name = "Group",
#                        col_name = "Q21")

```

```{r}
#| label: tbl-q22
#| tbl-cap: "Q22. Confidence in Fair Voting Process by Experiment Condition in Maricopa County, AZ by Experiment Condition"
#| tbl-cap-location: top

# Crosstab of Q22 by treatment group
df |> 
  janitor::tabyl(group, q22, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q22') |>
  gt::gt(
    rowname_col = 'Q22',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q22') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Q22. Confidence in Fair Voting Process for Maricopa County, AZ"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q22.How confident are you that the voting process will be fair in Maricopa County, AZ?",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# df |> 
#   janitor::tabyl(group, q22, show_na = F) |>  
#   janitor::adorn_totals('both', na.rm = T) |> 
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |> 
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'Q22') |> 
#   kableExtra::kbl() |> 
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}

df |> 
  janitor::tabyl(q7, q22, group, show_na = F) |> 
  # janitor::adorn_totals("both") |>
  janitor::adorn_percentages("row") |> 
  janitor::adorn_pct_formatting(digits = 2, affix_sign = F) |>
  janitor::adorn_ns("rear") |> 
  janitor::adorn_title("combined", 
                       row_name = "Q7. Legitimacy of 2020 Election",
                       col_name = "Q22") |>
  bind_rows(.id = "group_var") |>   
  gt::gt(
    rowname_col = "row_var",
    groupname_col = "group_var") |>  
  gt::row_group_order(groups = c("Control", "Treatment")) |> 
  gt::cols_label(
    c('Q7. Legitimacy of 2020 Election/Q22') ~ "Q7. Legitimacy of 2020 Election") |>
  # gt::cols_label_with(columns = '...1', fn = ~sjlabelled::get_label(df$q5.clps)) |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 3:6,
    label = "Q22. Confidence in Fair Voting process for Maricopa County, AZ"
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q22.How confident are you that the voting process will be fair in Maricopa County, AZ?",
    locations = gt::cells_column_spanners()) |> 
  gt::tab_footnote(
    footnote = "Q7.Regardless of whom you supported in the 2020 election, \ndo you think Joe Biden's election as president was legitimate, or was he not legitimately elected?",
    locations = gt::cells_column_labels('Q7. Legitimacy of 2020 Election/Q22')
    ) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )
```

```{r}
#| label: tbl-q23
#| tbl-cap: "q23. Confidence in Fair Voting Process by Experiment Condition in Maricopa County, AZ by Experiment Condition"
#| tbl-cap-location: top

# Crosstab of q23 by treatment group
df |> 
  janitor::tabyl(group, q23, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q23') |>
  gt::gt(
    rowname_col = 'Q23',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q23') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Q23. Confidence in Fair Voting Outcomesfor Maricopa County, AZ"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q23.How confident are you that the voting outcomes will be fair in Maricopa County, AZ?",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# df |> 
#   janitor::tabyl(group, q23, show_na = F) |>  
#   janitor::adorn_totals('both', na.rm = T) |> 
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |> 
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'q23') |> 
#   kableExtra::kbl() |> 
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}

df |> 
  janitor::tabyl(q7, q23, group, show_na = F) |> 
  # janitor::adorn_totals("both") |> 
  janitor::adorn_percentages("row") |> 
  janitor::adorn_pct_formatting(digits = 2, affix_sign = F) |>
  janitor::adorn_ns("rear") |> 
  janitor::adorn_title("combined", 
                       row_name = "Q7. Legitimacy of 2020 Election",
                       col_name = "Q23") |>
  bind_rows(.id = "group_var") |>   
  gt::gt(
    rowname_col = "row_var",
    groupname_col = "group_var") |>  
  gt::row_group_order(groups = c("Control", "Treatment")) |> 
  gt::cols_label(
    c('Q7. Legitimacy of 2020 Election/Q23') ~ "Q7. Legitimacy of 2020 Election") |>
  # gt::cols_label_with(columns = '...1', fn = ~sjlabelled::get_label(df$q5.clps)) |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 3:6,
    label = "Q23. Confidence in Fair Voting Outcomesfor Maricopa County, AZ"
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q23.How confident are you that the voting outcomes will be fair in Maricopa County, AZ?",
    locations = gt::cells_column_spanners()) |> 
  gt::tab_footnote(
    footnote = "Q7.Regardless of whom you supported in the 2020 election, \ndo you think Joe Biden's election as president was legitimate, or was he not legitimately elected?",
    locations = gt::cells_column_labels('Q7. Legitimacy of 2020 Election/Q23')
    ) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )
```

```{r}
#| label: tbl-q24-crab
#| tbl-cap: "q24. Confidence in Fair Voting Process by Experiment Condition in Maricopa County, AZ by Experiment Condition"
#| tbl-cap-location: top

# Crosstab of q24 by treatment group
df |> 
  janitor::tabyl(group, q24, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q24') |>
  gt::gt(
    rowname_col = 'Q24',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q24') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Q24. Confidence in Fair Voting Outcomesfor Maricopa County, AZ"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q24.How confident are you that the voting outcomes will be fair in Maricopa County, AZ?",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# df |>
#   janitor::tabyl(group, q24, show_na = F) |>
#   janitor::adorn_totals('both', na.rm = T) |>
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |>
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'q24') |>
#   kableExtra::kbl() |>
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}
#| label: tbl-q25-crab
#| tbl-cap: "Concern for violence, threats, or voter intimidation in Maricopa County, AZ by Experiment Condition"
#| tbl-cap-location: top

# Crosstab of Q25 by treatment group
df |> 
  janitor::tabyl(group, q25, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q25') |>
  gt::gt(
    rowname_col = 'Q25',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q25') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Concern for violence, threats, or voter intimidation in Maricopa County, AZ"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q25. Thinking about Maricopa County, AZ, how concerned should voters feel \nabout potential violence, \nthreats of violence, or intimidation while voting in person at their local polling place?",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# alternative crosstab
# df |> 
#   janitor::tabyl(group, q25, show_na = F) |>  
#   janitor::adorn_totals('both', na.rm = T) |> 
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |> 
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'Q25') |> 
#   kableExtra::kbl() |> 
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}
#| label: tbl-q26-crab
#| tbl-cap: "Q26. In-person voter safety in Maricopa County, AZ by Experiment Condition"
#| tbl-cap-location: top

# Crosstab of Q26 by treatment group
df |> 
  janitor::tabyl(group, q26, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q26') |>
  gt::gt(
    rowname_col = 'Q26',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q26') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Confidence for In-person voter safety in Maricopa County, AZ"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q26. How confident, if at all, are you that in person polling places in Maricopa County, AZ \nwill be safe places for voters to cast their ballots during the upcoming elections in November?",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# df |> 
#   janitor::tabyl(group, q26, show_na = F) |>  
#   janitor::adorn_totals('both', na.rm = T) |> 
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |> 
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'Q26') |> 
#   kableExtra::kbl() |> 
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}
#| label: tbl-q27-crab
#| tbl-cap: "Q27. Approval for Election official in Maricopa County, AZ"
#| tbl-cap-location: top

df |> 
  janitor::tabyl(group, q27, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q27') |>
  gt::gt(
    rowname_col = 'Q27',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q27') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Approval for Election official in Maricopa County, AZ"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q27. Do you approve or disapprove of the way election officials in Maricopa County, AZ are handling their jobs?",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# df |> 
#   janitor::tabyl(group, q27, show_na = F) |>  
#   janitor::adorn_totals('both', na.rm = T) |> 
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |> 
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'Q27') |> 
#   kableExtra::kbl() |> 
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}
#| label: tbl-q29-crab
#| tbl-cap: "Q29. Should Local Area Adopt Recruitment Program of Maricopa County, AZ"
#| tbl-cap-location: top

df |> 
  janitor::tabyl(group, q29, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q29') |>
  gt::gt(
    rowname_col = 'Q29',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q29') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Should Local Community Adopt Maricopa County, AZ Recruitment Program"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q29. How much would you like to see your local community adopt a program for recruiting elections staff like the one you read about in Maricopa County, AZ?",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# df |> 
#   janitor::tabyl(group, q29, show_na = F) |>  
#   janitor::adorn_totals('both', na.rm = T) |> 
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |> 
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'Q29') |> 
#   kableExtra::kbl() |> 
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}

df |> 
  janitor::tabyl(group, q41.4, show_na = F) |>
  janitor::adorn_percentages(denominator = 'row') |>
  janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
  janitor::adorn_ns() |> 
  janitor::adorn_title(
    'combined',
    row_name = "Group",
    col_name = 'Q41.4') |>
  gt::gt(
    rowname_col = 'Q41.4',
    groupname_col = 'Group' 
  ) |>  
  gt::cols_label(
    c('Group/Q41.4') ~ "Group") |> 
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |> 
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_labels()
  ) |> 
  gt::tab_spanner(
    columns = 2:5,
    label = "Election staff and volunteers include military veterans and their family members"
  ) |>  
  gt::tab_style(
    style = gt::cell_text(weight = 'bold'),
    locations = gt::cells_column_spanners()
    ) |> 
  gt::tab_footnote(
    footnote = "Table reflects row percentages. NAs omitted") |> 
  gt::tab_footnote(
    footnote = "Q41.4. How would the following impact your confidence in the fairness and accuracy of elections conducted this November? Election staff and volunteers include military veterans and their family members from the community",
    locations = gt::cells_column_spanners()) |> 
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

# "Regardless of whether any of these are actually the case, how would the following impact your confidence in the fairness and accuracy of elections conducted this November?"
# Election staff and volunteers include military veterans and their family members from the community.

# df |> 
#   janitor::tabyl(group, q29, show_na = F) |>  
#   janitor::adorn_totals('both', na.rm = T) |> 
#   janitor::adorn_percentages(denominator = 'row') |>
#   janitor::adorn_pct_formatting(digits = 1, affix_sign = F) |>
#   janitor::adorn_ns() |> 
#   janitor::adorn_title(
#     'combined',
#     row_name = "Group",
#     col_name = 'Q29') |> 
#   kableExtra::kbl() |> 
#   kableExtra::kable_styling(
#     bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
#     latex_options = "basic",
#     font_size = 13
#   )

```

```{r}
#| label: tbl-7
#| tbl-cap: "Confidence Impact on In-person Voter Safety"
#| results: asis

# Simple Crosstab to show response breakdown by experiment condition
# NAs omitted. 
df |>
  mutate(
    confimpct_vets = dplyr::case_when(
      q43.4.clps == "increase" ~ 'Increase Confidence',
      q43.4.clps == "no_impact" ~ 'No Impact',
      q43.4.clps == 'decrease' ~ 'Decrease Confidence'
      )) |> 
  janitor::tabyl(group, confimpct_vets,  show_na = F) |> 
  janitor::adorn_totals('both') |> 
  janitor::adorn_percentages('row') |> 
  janitor::adorn_pct_formatting(digits = 2, affix_sign = T) |> 
  janitor::adorn_ns("rear") |> 
  janitor::adorn_title(
    'combined',
    row_name = 'Condition',
    col_name = 'Election staff include veterans and family') |>
  dplyr::relocate(
    "No Impact", .before = "Increase Confidence"
  ) |> 
  kableExtra::kbl() |> 
  kableExtra::footnote(
    general = "NA omitted. % (n)"
    ) |> 
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "bordered", "condensed", "responsive"),
    latex_options = "basic",
    font_size = 13,
    full_width = T,
    fixed_thead = T
  )

```