### 2025-02-26 index.qmd

---
title: title
subtitle: subtitle
shorttitle: A Short title # aog-article-pdf
shortauthors: Isaiah      # aog-article-pdf
authors:
  - name: Isaiah Espinoza
    email: gespinoz@umd.edu
    roles: author
    affiliation:
      - name: University of Maryland
        department: Department of Government and Politics
        city: College Park
        state: MD
        country: US
reference-section-title: References
date: "2024-10-16"
date-modified: today
knitr: 
  opts_chunk: 
    dev: "ragg_png" # same as knitr::opts_chunk$set(dev = "ragg_png")
---

```{r}
#| label: setup
#| include: false

set.seed(1234)

# load packages
library(tidyverse)
library(easystats)

options("huxtable.latex_use_fontspec" = TRUE) # if TRUE, uses the LaTeX “fontspec” package which lets you use the same fonts in TeX and HTML. You will need to use the xetex or xelatex engine for output.

# load data
load("data/research-paper-df-20250203.Rdata") #df

# load data dictionary/codebook
codebook <- read.csv("data/survey_codebook.csv")
df_dict <- read.csv("data/df_dictionary.csv")

```

\pagebreak

## Introduction

Election administration officials make efforts to sustain public trust and confidence in the fairness and accuracy of elections, and attempt to boost such confidence where it may be deprived. Concerns for safety have developed among election staff and voters in more recent elections. Regular measures are taken to enhance the *trustworthiness*[^1] of the electoral process through practices meant to improve the conduct, transparency, or overall administration of elections in the United States.

Although election officials undertake great efforts to enhance the trustworthiness of election administration, public *trust* in elections is a psychological construct influenced by many things outside of election official control such as partisanship or elite rhetoric [@hooghe2018; @sances2015]. Moreover, a person's evaluation of the election in hindsight is often influenced by the election outcome itself [@daniller2019; @stewart2022]. Thus, measures taken by election officials can be undermined, trivialized, or made irrelevant depending on how one feels after the election results have come out.

Such volatile attitudes and evaluations post-election can leave a lasting impression that election officials must contend with upon the next election cycle [@levendusky2024; @bowler2024]. For instance, we have witnessed many people's outright refusal to accept the 2020 U.S. election results as legitimate despite consistent review of the evidence confirming the results as fair and accurate. Such a case demonstrates that public trust in elections is, at best, only partial to trustworthiness of election administration in the United States[^2].

Even though election officials can do a lot to secure election integrity, there's not much they can do to cement public confidence after election night passes. At best, election officials can ease public insecurities prior to election night.

One point of contention that election officials have faced in the past regard evaluation of election workers. Previous literature has focused on how voter interaction with election workers [@claassen2008], or the voter experience generally [@atkeson2007], influences evaluations of election administration. As such, election worker competency has been examined as a factor significant to evaluations of performance of elections [@hall2007; @hall2009]. However, considering that individual perceptions and preconceived notions play a huge role in cognition [@cikara2014; @vanbavel2021], it is reasonable to expect that the group an election worker hails from would be an important influence upon the voter's evaluation of the electoral process.

Supposing such is the case, we can expect that information about *who* (i.e., which groups) election officials are targeting in publicized recruitment efforts would lessen particular election insecurity, and concurrently, boost confidence. That is to say, it is reasonable to expect that telling people *who* will be working and volunteering as election staff would ease election insecurity, and therefore improve confidence that the election will be conducted fairly, accurately, and safe for all involved.

In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment support the notion that emphasizing veterans as the target of election worker recruitment efforts eases pre-election insecurity to a limited extent. Expectations of electoral fraud and concerns for voter safety were lower among those who read a fabricated announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate.

This paper is structured as follows. First, I provide a brief background on public trust in election administration. I synthesize a review of relevant literature with a focus on how political and social science has conceptualized and ascertained public trust in elections. Next, I supply reasoning for why military veterans are singled out as the relevant subset of the population in this study. I then explicitly provide the simple theory and testable hypotheses of the study before moving on to describe the design of the survey, the measurement instruments therein, and the conceptual and operational definitions of the variables of interest. I explain my reasoning and method for constructing the primary dependent variable, which I broadly refer to as *confidence in elections*. I preview the different methods of analysis alongside presentation of the results. Before concluding, I dedicate time to discuss the limitations of the study and lessons learned. I close by offering my interpretation of the results and suggest potential avenues for future inquiry.

## Background: Election Administration and Public Confidence

Election officials have tried hard to inspire confidence in the administration and conduct of elections by improving the degree to which elections are trustworthy. Development and implementation of procedures such as post-election auditing of ballots and logic-and-accuracy testing of ballot tabulation equipment are prominent examples adding to the long history of efforts to enhance the trustworthiness of election administration in the United States.

Prior to the year 2000, one of the main issues facing election administration was recruiting enough election workers to volunteer at the polls (i.e., poll workers) [@maidenberg1996]. Election worker recruitment is still much of an issue in the current era as it was then, perhaps worse [@ferrer2024]. In addition to ensuring election admin offices were adequately staffed, the controversy of the 2000 general election made the public more attentive to issues concerning the conduct and administration of elections. In particular, voting technology [@herrnson2009] and election worker competence was a of interest in election studies [@claassen2008; @hall2009; @hall2007]. Following the passage of the Help America Vote Act in 2002, election officials efforts to boost public confidence in the conduct and administration of elections revolved primarily around the accuracy of vote counts, ballot tabulation equipment or voting machines, the commitment of election staff, and more [@atkeson2007].

In 2024, election officials made valiant efforts to boost public confidence elections within an intensified political climate that appeared quite hostile to election officials [@edlin2024; @brennancenterforjustice2024]. Although polling around the time indicated that most people thought that U.S. elections would be run at least somewhat well [@nadeem2024], many election officials nationwide took efforts to assuage the worry of those most skeptical.

Election anxiety was high in the lead up to the 2024 elections in the United States. Concerns for voter safety and the prospect of political violence remained prescient and compelled many local officials to prepare for the worst [@edlin2024; @doubek2024]. Election officials in Washoe County, Nevada, installed panic buttons for election staff that would alert a monitoring center to summon law enforcement [@lincoln2024]. Nevada also passed a law making it a felony to harass, threaten, or intimidate election workers [@nevadasecretaryofstate2023]. Leading up to election day, news outlets reported that election work had become a seemingly dangerous job [@wire2024]. A Brennan Center survey report stated that, "...large numbers of election officials report having experienced threats, abuse, or harassment for doing their jobs" [@edlin2024]. Concerns over the fairness of elections and accuracy of vote counts intensified, heightening concerns over the prospect of political violence and, in turn, increased worry for the safety of voters and election workers alike.

Suffice to say, pre-election anxiety consists of more than confidence in fairness and accuracy of vote counts in light of added safety concerns. It is not hard to recognize that increased tension in the pre-election period makes for a volatile political environment. Sustaining trustworthiness in election administration is only more difficult in an environment where turnover of election workers increases and the struggle to recruit volunteers worsens in light of safety concerns. Since election worker performance is significant to public evaluation of elections, added safety concerns that drive out election staff and repel volunteers can only detract from trustworthiness of the institution.

## Literature Reivew

Trust and confidence in the conduct of elections concerns aspects of elections that fall squarely within the institution of election administration. At this level, for instance, public trust is ascertained by capturing assessments about the perceived accuracy of vote counts (e.g., whether votes are/were counted as intended). Or in other words, by the extent the public is confident in the accurate administration of elections.

Assessing the public's *trust* in elections has not been straightforward, however. Inquiry into public trust in elections has been approached by scholars of political science in many different ways [@cook2005; @hardin2004], often distinguishable by the scope of the research question and more or less constrained by the particular conception of public trust. Quite often, trust in election administration is conflated with trust in government *writ* large, government legitimacy, government or system responsiveness, or even satisfaction with democracy [@daniller2019]. At this level, not only is the level of public trust in elections sometimes vague, but there's little consideration over the difference between such attitudes pre-election and post-election. In contrast, a considerable amount of research tends to conceive of public trust in accordance with the institution in question [@atkeson2007; @hooghe2018].

Elections administration is just one institution a part of the larger set of institutions which form the electoral system. As such, the performance of the institution along with the rest "...lends credibility to the outcome of an election: whether it is considered by citizens and the international community to be fair and legitimate." [@stewart2022, 236]. Intuitively, enhancing public trust in elections would best be accomplished by enhancing the *trustworthiness* of the institution, i.e., consistently doing the things that election officials already regularly do come election time. However, trust and confidence in elections has become ever more precarious over the last few election cycles. Especially considering public polling data since 2000 shows that confidence that votes were, or would be, counted as intended was in a consistent decline despite efforts towards bolstering election integrity and trustworthiness [@sances2015]. This is even more pronounced considering the role that partisanship has had on such confidence over accuracy of vote count [@sances2015; @stewart2022].

There's also stark difference in public trust before the election has occurred compared to after, a phenomenon referred to as the "winner-loser gap"; the "winners" are those who supported the winning candidate and the "losers" are those who supported the losing candidate. Much research has been dedicated to analyzing the sentiment of electoral winners vs losers, and vice versa [@nadeau1993; @daniller2019]. Opinions of electoral trust gathered after the election has occurred are limited considering the well-recognized impact that the electoral outcome itself has on feelings of public trust in elections [@daniller2019].

As such, it is questionable whether we can characterize public trust in the pre-election period as the same trust after the election results have come out. The former is *anticipatory*---i.e., the kind that is more or less anxious given the uncertainties surrounding the election. The latter is *empirical*---a judgement discerned in hindsight after the experience of the election event has occurred. Evaluations of one's own trust and confidence in election administration based on the experience of voting is influenced by that very experience. As discussed by @stewart2022, not only was confidence in vote count accuracy influenced by the voter experience, but so too were evaluations of election officials [-@stewart2022, 242-243]. Not to mention the influence that the election results would also have on such evaluations. In other words, evaluations of trust and confidence post-election are influenced by one's interpretation of their experience. The subtle difference is simply the degree of uncertainty one feels in anticipation of the next election event. This study focuses primarily on that *anticipatory* kind of confidence, which speaks more to those insecurities[^3] based on perceptions of the institution's trustworthiness than upon the particular voter experience.

Regardless of the measures taken by election officials to boost public confidence in the *trustworthiness* of election administration (e.g., conducting audits, testing election machines), public *trust* and confidence in elections more generally is apt to shift dramatically post-election based on factors such as partisanship, elite rhetoric, particular state policies, and more [@nadeau1993; @carter2024; @coll2024a]. Moreover, prior research has found that evaluation of election workers themselves are an important factor when it comes to levels of public confidence in the electoral process [@hall2007; @hall2009; @claassen2008]. Such studies focused on the quality of the voter experience with reference to the interaction between voter and election worker.

Beyond the general competence of election workers, however, the quality of the voter experience may be influenced merely by *a priori* impressions about *who* comprises election staff and volunteers. Political and other social science researchers have recognized for some time the power that group identity can have over attitudes and perception [@xiao2016; @xiao2012; @vanbavel2021]. As such, we can expect that information identifying the particular groups being recruited to serve as election staff will be enough to improve trust and confidence in election administration, and lessen expectations of electoral fraud. The next section elaborates on why military veterans are of particular interest in this regard.

## Why Veterans

Election officials are likely to be agnostic as to who dedicates their time to civil service such as election work. Staffing issues have been an issue since at least the 1990s [@maidenberg1996; @ferrer2024]. In 2020, such efforts were made far more difficult by the COVID-19 pandemic [@mena2020; @abbate2020a]. Sure, it makes sense to recruit veterans, but no more than any other group. After all, ensuring election offices are adequately staffed is everyone's problem. So there's no special reason to target veterans for recruitment above other groups. Indeed, there's no reason to discriminate recruitment efforts at all if the point is purely to fill staffing vacancies. However the interest in veterans as a special group to consider arose in light of increased and sudden efforts to recruit veterans into election work.

The reason why veterans arose as a special subset of the population to consider for this inquiry is because there was a sudden push to target veterans for recruitment efforts that arose shortly after the events on Capitol Hill on January 6th, 2021. After the 2020 election, large efforts were made to recruit military veterans and their families to work or volunteer as election staff [@wetheveterans2022; @nflfootballoperations2022]. Prior to that point, young people were sometimes given special mention as targets of election worker recruitment efforts [@powerthepolls2020; @herndon2020]. Generally, however, recruitment efforts cast a wide net, indiscriminate of who applies [@ross2020; @conde2020].

One can speculate that the motivation to associate military veterans with civic engagement and democracy may be intended to counter negative perceptions and impressions given by the proportion of veteran service members arrested for taking part in the events on January 6th [@milton2021; @loewenson2023; @jensen2022a]. Especially with regard to research demonstrating that willingness to support violent efforts to overturn election results (in support of Trump) is, on average, more common among veterans than among matched samples of non-veterans [@pape2024]. This is in addition to a strengthened association portrayed in media outlets between military veterans and militias [@steinhauer2020]. Prior research has substantiated such a connection between veterans and militia groups. @cooter2024 notes from her 3-year ethnographic fieldwork among Michigan militia members that, "...approximately 40% of militia leaders and 30% of members had previous military experience. Most of these veterans actively sought out such groups, as opposed to being recruited by them" [-@cooter2024; see also @cooter2013]. Thus, countering such associations by promoting a different image of veterans to the mass public and veterans alike seems like a reasonable motivation. Yet such speculation is just that.

That being said, the general public perception, attitudes, or even stereotypes about military veterans are significant to consider. Although military recruitment shows a downward trend as of late, public perception of veterans are overwhelmingly positive [@kleykamp2023]. Veterans are a particularly potent group where mention of veteran status seems to have a positive, calming, or nullifying effect on attitudes. For instance, recent research shows that, during his campaign in the 2020 Democratic Primaries, Pete Buttigieg's military background mitigated discrimination against him when he was presented as a veteran married to man [@magni2024]. Similar research has also found that a candidate's veteran status affords them better evaluations regarding competency in particular issue areas (e.g., war competence) [@hardy2019]. Moreover, veteran status seems to mitigate or nullify usual stigmas associated with mental illness. That is to say, there is negative stigma associated with mental illness [@corrigan2002] and such stigma incurs labor market discrimination [@hipes2016], but evidence suggests that veteran status overrides such stigma and discrimination [@maclean2014]. Or, in another light, mental illness is seemingly more *understandable* (i.e., permissible) for veterans given the presumptive reasons for their mental strife. And media framing as such plays an important role on public perception [@kleykamp2015].

## Theory and Hypotheses

Considering the generally positive perception afforded to veterans by the public, it is reasonable to suspect that efforts that promote recruitment of veterans and their family members to work as election staff and volunteers would boost public confidence in election administration. Indeed, that is the primary hypothesis of this study. Although citizens may vote in various ways across the country (e.g., by mail, ballot drop box, in-person), the simple announcement that election officials are engaging in efforts to recruit veterans to work as election staff may boost confidence in elections administration regardless of how, or whether, an individual plans to vote. Formally,

> H~1~: Announcements that election officials are recruiting miilitary service veterans to work as election staff and volunteers will be associated with greater confidence in elections administration compared to announcements that do not mention military veterans.

I considered that veterans would elicit high degrees of approval and support for whatever cause or issue presented. For example, I'd expect high approval and support for recruiting veterans to baseball teams just as much as for recruiting them to work as election staff. Likewise, asking participants whether they supported programs intent on recruiting anyone to work as election staff would garner support regardless of the target group. Considering that support and admiration for veterans is generally high among the population [@kleykamp2023], then discerning the impact on one's confidence would require a survey experiment designed to determine whether publicized recruitment efforts targeting veterans as a group would have any special effect on confidence in elections administration.

Since the results and events following the 2020 election loomed large in anticipation of the U.S. 2024 general elections, it is reasonable to take into account beliefs about legitimacy of the 2020 election. This group who denies the legitimacy of the 2020 election results were possibly the most distrusting of the 2024 election. However, given that veteran service members are held in high regard, generally speaking, announcing that veterans are being actively recruited to work as election staff and volunteers is expected to positively influence confidence in elections among those who remained firm in the belief that the 2020 election results were illegitimate. I take into account that legitimacy beliefs about the 2020 election and expectations of electoral fraud are related to partisanship and include partisanship as a control in the analysis.

In addition, I expect that both expectations of electoral fraud and concerns for voter safety would lessen (shrink) when presented with information that veterans are being actively recruited to work and volunteer in election offices.

Finally, prior research has shown that confidence that one's own vote would be counted as intended has been much stronger than confidence that votes nationwide would be treated likewise [@sances2015; @stewart2022]. In other words, the public is more trusting of elections closer to home. Taking into account that confidence is lower for elections conducted outside of one's local area, then information that veterans are being recruited to help conduct elections in locations beyond one's locale may close this gap between confidence in elections within one's local area and confidence in elections elsewhere. Although it may be difficult to discern whether recruiting veteran service members would improve confidence in elections near one's local area in light of such hometown favoritism, a positive effect on confidence in elections held beyond one's local area---where confidence is already expectedly lower---can help discern the impact of the group alone. Therefore, any observed disparity of confidence in elections (i.e., the "confidence gap") between elections within one's local area and elections beyond will be smaller when people are presented with information about election official recruitment efforts of veteran service members.

## Experiment Design and Survey Measures

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve confidence in elections and ease insecurity, a recent experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment ($n =$ 650) control ($n =$ 637) conditions[^4].

Survey participants read either a treatment or control vignette, which was a fabricated news article about efforts in Maricopa County, AZ to recruit election staff and volunteers for the 2024 general election. The treatment vignette referred to a program designed to recruit veterans and their family members and describes an interviewee "Jordan Braxton" as an Army veteran. The control vignette simply omitted any mention of veterans and their family members, and didn't describe "Jordan Braxton" as an Army veteran. Beyond those small differences and the headlines, the article vignettes are identical. Therefore, effects can be attributed to the information about veterans in the treatment vignette[^5].

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignette due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. Because of this, treatment effects are potentially limited or constrained to attitudes concerning the location specified in the vignette. A block of survey items asked specifically about Maricopa County, AZ, followed by an identical block of items that asked the same questions about one's local area. Later, I compare and discuss results between the items pertaining to different settings.

Over the past two decades it became commonplace for national polls to gauge public confidence in election administration (i.e., voter confidence) by asking some variety of the question, "How confident are you that your vote \[will be/was\] counted as you intended in the most recent election?" [@hall2009; @sances2015; @stewart2022]. In addition, since 2008, the Survey of the Performance of American Elections (SPAE) has included a good number of relevant questions to more thoroughly assess trust and confidence in election administration. Such questions inquire into the voter experience with the institution more directly. This study borrows, modifies, or takes inspiration from certain question items found within the 2022 SPAE and other survey items from the Pew Research Center's 2018 American Trends Panel wave 38 [@dunn2018].

Note that, except for the item assessing voter confidence (i.e., votes counted as voters intend), most other question items from the 2022 SPAE inquire into the voter experience after the election has occurred. The items in this study, however, all inquire into confidence over one's expectations in anticipation of upcoming elections. In addition, other survey questionnaires ask similar questions that specifically contrast between a respondent's own local community, their state, and nationwide; this is in addition, or complementary to, items that distinguish between local area officials, state election officials, state government, and the U.S. federal government. The survey of this study, instead, focused on attitudes regarding election workers such as election officials, staff, and volunteers specific to the location mentioned in the vignette---Maricopa County, AZ---as well as specific to one's local area[^6].

The survey included items that assessed both trust in elections administration and expectations of electoral fraud, referred to simply as *trust* and *distrust*, respectively. Responses from these two sets of items were utilized to construct the primary dependent variable, *confidence in elections*.

Prior to the reading the treatment or control vignette, participants responded to distinct survey items assessing one's favorability to local election officials ("In general, how favorable or unfavorable is your impression of local election officials?") and legitimacy of the 2020 election results ("Regardless of whom you supported in the 2020 election, do you think Joe Biden's election as president was legitimate, or was he not legitimately elected?"). The former presented five response options from "Strongly favorable" to "Strongly unfavorable"; the latter item presented a binary of either "Legitimate" or "Not Legitimate".

The conventional multiple item assessment of partisanship was included and used to construct a variable of partisanship consisting of three categories: Democrat, Republican, and Independent. In this case, Independents are "true" Independents in that they do not identify with either political party and do not "lean" toward either Republican or Democratic parties. Self-described "Independent leaners" are re-coded as falling into the partisan category in which they lean e.g., Independents who expressly report that they lean Republican are denoted as Republican.

Finally, a pair of survey items assessed concerns for potential violence and confidence in voter safety. The pair was asked once in regard to Maricopa County, AZ and again with respect to one's local area[^7].

### Trust and Distrust in Elections

A set of five survey items measured public trust in elections by inquiring into the degree of confidence that votes will be counted as voters intend, confidence that electoral systems are secure from technological threats, perceived commitment of election staff, confidence that outcomes will be fair, and confidence that the voting process will be fair. Each item in this series presented four response options from "Not at all \[confident/committed\]", "Not too \[confident/committed\]", "Somewhat \[confident/committed\]", and "Very \[confident/committed\]". In total, five survey items measure different aspects of trust in elections administration. The items measuring confidence in elections include:

Another series of five survey items captured an individual's level of *distrust* in elections administration based on the extent to which they *expect electoral fraud* to occur. These items were prefaced with the question, "How likely do you think any or all of the following will happen during this year´s elections in \[Maricopa County, AZ/ your local area\]?" Each item presented four response options: "Not likely at all", "Not too likely", "Somewhat likely", and "Very likely".

```{r}
#| label: tbl-items
#| tbl-cap: "Survey Items for Trust and Distrust in Elections"

trust_txt <- df |> select(q19, q20, q21, q22, q24) |>
  labelled::set_variable_labels(
    q21 = "How commited do you think election staff and volunteers in Maricopa County, AZ will be to making sure the elections are fair and accurate?"
  ) |> 
  labelled::generate_dictionary() |> 
  dplyr::as_tibble() |> 
  select(Trust=label)

distrust_txt <- df |> select(q28_1:q28_5) |>
  labelled::generate_dictionary() |> 
  dplyr::as_tibble() |> 
  select(Distrust=label)

conf_items_text <- trust_txt |> 
  mutate(Distrust = pull(distrust_txt))

kableExtra::kbl(conf_items_text, booktabs = T,
                col.names = c("Trust", "Distrust"),
                linesep = "\\addlinespace") |> 
  kableExtra::kable_styling(full_width = T,
                            font_size = 8,
                            latex_options = c("HOLD_position")) |> 
  kableExtra::column_spec(column = 1:2, width = "8cm")

```

Similar to the items capturing trust in elections (@tbl-items), I computed a summated score for each respondent reflecting the extent of their distrust in elections.

Upon construction of the two summated scales of *trust* and *distrust*, internal consistency reliability coefficient $\alpha$ were estimated for both and found to be adequate ($\alpha$ \> 0.8); the trust in elections scale $\alpha$ = 0.93 and the distrust scale $\alpha$ = 0.86. In other words, the items in each of the scales were intercorrelated well enough for each to compose a single scale presuming a unidimensional construct for each.

I expected item responses from these two scales to be inversely correlated among the sample, but not mutually exclusive. Indeed, this is what I find. Polychoric correlations between the trust and distrust items negatively correlate as expected\[\^See Appendix C\], as does the correlation between the two sum score scales. Given the ordinal nature of the variable items, I conducted a Spearman's rank correlation test [@spearman1907] and found a negative correlation of $\rho = -0.49$ ($95\%$ CI \[-0.53, -0.44\]; Kendall's $\tau = -0.39$) between scores on the two scales for the AZ items (For the local item scale score correlation: $\rho = -0.52$, $95\%$ CI \[-0.56, -0.48\]; Kendall's $\tau = -0.39$). The negative correlation between the trust and distrust items and scores makes intuitive sense.

## Data Specification

The primary outcome of interest concerns public confidence in elections administration compensated by one's distrust in elections, i.e., their expectations of electoral fraud. Because trust and distrust are assessed by multiple Likert items, conventional methods call for combining the items of each set into separate Likert scales (also known as summated rating scales). Since the item response options are not symmetric---i.e., not bipolar with contrasting positive and negative response options to include a neutral middle option---then the *trust in elections* scale could only be understood as a measure of trust in the positive use of the term (i.e., unipolar). Likewise, the expectation of fraud items fit to a scale that corresponds to a unipolar measure of confident *distrust*, so to speak. However, many of the items comprising the trust scale directly oppose items capturing distrust. For instance, one item from the trust scale has the respondent report their confidence that votes will be counted as voter's intend, whereas a separate item from the distrust scale has the the respondent report their expectation that many votes will not be counted. Positive responses to these two items directly contradict each other, and would thus cancel out.

The relationship is that trust and distrust represent opposing degrees of confidence along the same spectrum. However, despite the inherent contrast, one can hold both positive degrees of trust and distrust, feeling similarly confident in both respects. Accordingly, equivalent degrees of trust and distrust cancel out, rendering one relatively more *insecure* about their expectations of the future despite the fact that they may feel confident in either direction. When trust and distrust are considered as reflective of one's confidence about future expectations, then both can be placed along the same spectrum. Therefore, a lack of confidence denotes insecurity.

Something may cause a person to have greater expectations that election fraud will occur, which in turn will always lower their trust in elections administration. In contrast, a person may feel a bit more confident in the integrity of elections, but this confidence is constrained by the extent to which they expect election fraud to occur. This means that a person's baseline level of confidence in elections admin must take into account their present degree of distrust---i.e., their expectations of electoral fraud.

Those who are, or are close to being, equally confident in both directions would be the most *insecure* about their expectations. Having no trust in elections and no expectation that electoral fraud will occur is functionally equivalent to having equal degree of both. Quite literally, this reflects an uncertain feeling of "it could go one way or the other" with respect to the integrity of elections administration. Operationally, this would reflect a middling point on a scale of *confidence*.

In light of this, I reverse coded positive expectations of electoral fraud and added responses from the distrust scale to the trust in elections scale. That is, item responses on the trust scale ranged from 0 ("Not confident at all") to 3 ("Very confident), whereas responses on the distrust scale ranged from 0 ("Not likely at all") to -3 ("Very likely"). Combining responses in this way resulted in an overall measure of *confidence in elections* to serve as the dependent variable. On this scale, positive values reflect higher confidence in elections, negative values reflect greater distrust, and values closer to zero correspond to an individual's degree of insecurity[^8].

The advantage is that this permits a better test of the treatment effect as theorized. Of course, it is easy enough to examine the effect of the treatment on both trust and distrust separately. However, the theory presents a hypothesis that the treatment will impose influence upon both variables *concurrently and in a positive direction* in light of the relationship between trust and distrust. Confidence increases insofar as trust in elections increases and distrust decreases; since distrust detracts from trust, then an accurate measure of confidence must take that constraint into account.

## Results

I first review responses to the survey items capturing trust and distrust, respectively. When distinguished by experiment condition, trust in elections admin is noticeably higher for those who read the treatment vignette. The difference is more pronounced for some items over others. That being said, there is a definite distinction between items pertaining to Maricopa County, AZ compared to items that pertain to a respondent's local area. In fact, the treatment appears to have hardly made a difference when the items concerned one's local area.

```{r}
#| label: fig-trustlikert
#| fig-cap: "Trust in Election Administration by Experiment Condition"

# Likert plot of trust and confidence in election admin comparing treatment and
# control groups
trust.likert.treat.az <- df |>
  # relabel factor labels for plotting
  mutate(across(
    c(q19, q20, q21, q22, q23, q24, q30, q31, q32, q33, q34, q35),
    ~ pewmethods::fct_case_when(
      .x %in% c("Not at all confident", "Not at all committed") ~ "Not at all confident",
      .x %in% c("Not too confident", "Not too committed") ~ "Not too confident",
      .x %in% c("Somewhat confident", "Somewhat committed") ~ "Somewhat confident",
      .x %in% c("Very confident", "Very committed") ~ "Very confident",
      TRUE ~ NA
    )
  )) |>
  ggstats::gglikert(
    include = c(q19, q20, q21, q22, q24),
    y = 'group',
    facet_rows = vars(.question),
    symmetric = F,
    labels_size = 2.5,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Maricopa County, Arizona", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_blank()
  )

# confidence in elections between experiment conditions for items concerning local area
trust.likert.treat.lcl <- df |>
  # relabel factor labels for plotting
  mutate(across(
    c(q19, q20, q21, q22, q23, q24, q30, q31, q32, q33, q34, q35),
    ~ pewmethods::fct_case_when(
      .x %in% c("Not at all confident", "Not at all committed") ~ "Not at all confident",
      .x %in% c("Not too confident", "Not too committed") ~ "Not too confident",
      .x %in% c("Somewhat confident", "Somewhat committed") ~ "Somewhat confident",
      .x %in% c("Very confident", "Very committed") ~ "Very confident",
      TRUE ~ NA
    )
  )) |>
  ggstats::gglikert(
    include = c(q30, q31, q32, q33, q35),
    y = 'group',
    facet_rows = vars(.question),
    variable_labels = c(
      q30 = "Votes counted as intended",
      q31 = "Election officials will do good job",
      q32 = "Election workers will be committed",
      q33 = "Voting process will be fair",
      q35 = "Election technology will be secure"
    ),
    facet_label_wrap = 10,
    labels_size = 2.5,
    symmetric = F,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Local Area", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.y.left = element_blank(), # remove y-axis text on 2nd plot
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_text(angle = 0, size = 8)
  )

# trust.likert.treat.az
# trust.likert.treat.lcl

# Instead of two separate plots, arrange into one plot while remaining separated
# use `ggpubr` package to print multiple plots side-by-side (or alt. row-by-row)
ggpubr::ggarrange(
  trust.likert.treat.az,
  trust.likert.treat.lcl,
  ncol = 2,
  common.legend = T,
  legend = "bottom", 
  font.label = list(family = "Palatino Linotype")
)
```

Reviewing item responses distinguished by treatment condition reveals particulars of the treatment effect (@fig-trustlikert). Among the items that asked about Maricopa County, AZ, it is a bit easier to see upon which items the treatment vignette had the greatest effect. It appears that the main effect of the treatment is upon responses to the item that measures confidence in the security of election technology.

Regarding confidence in the commitment of election staff and volunteers, a higher proportion of respondents in the treatment group selected "Very confident" over "Somewhat confident" compared to the control group, but only for items concerning Maricopa County, AZ. Although confidence in election workers was generally high across the board, the shift from "Somewhat" to "Very" confident for items pertaining to election worker *commitment* to fairness and accuracy seems to clearly speak to the group being recruited, i.e., veterans and family members. However, the same visible shift in proportion from "Somewhat" to "Very" confident isn't observed when items asked about one's local area, which undermines the suggestion that the treatment effect can be cleanly attributed to respondent's feelings about veteran service members and their families. It may be the case that respondents harbor particular attitudes or sentiment about Maricopa County, AZ itself, or there may be nothing uniquely special about Maricopa County, AZ supposing that elections anywhere beyond one's local area generally garner slightly less confidence.

What this does suggest, however, is that trust in the fairness and accuracy of the electoral process is notably higher upon the notion that military veterans will be directly engaged as election staff and volunteers in places where such trust falters. The question turns to whether this also lowers distrust, or the extent to which one expects electoral fraud to occur.

```{r}
#| label: fig-distrustlikert
#| fig-cap: "Disrust in Election Administration by Experiment Condition"

# distrust in elections in AZ by experimental condition
distrust.likert.az <- df |>
  # reverse levels of factor vars to show positive difference in "not likely"
  mutate(across(c(q28_1:q28_5), ~forcats::fct_rev(.))) |> 
  ggstats::gglikert(
    include = c(q28_1:q28_5),
    y = 'group',
    facet_rows = vars(.question),
    variable_labels = c(
      q28_1 = "Voter fraud will occur",
      q28_2 = "Many votes won't be counted",
      q28_3 = "Many voters will be turned away",
      q28_4 = "Foreign Interference",
      q28_5 = "Election officials will discourage voting"),
    facet_label_wrap = 10,
    labels_size = 2.5,
    symmetric = F,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Maricopa County, Arizona", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_blank()
  )

# distrust in elections in Local area by experimental condition
distrust.likert.lcl <- df |>
  # reverse levels of factor vars to show positive difference in "not likely"
  mutate(across(c(q40_1:q40_5), ~forcats::fct_rev(.))) |> 
  ggstats::gglikert(
    include = c(q40_1:q40_5),
    y = 'group',
    facet_rows = vars(.question),
    variable_labels = c(
      q40_1 = "Voter Fraud will occur",
      q40_2 = "Many Votes won't be counted",
      q40_3 = "Many voters will be turned away",
      q40_4 = "Foreign Interference",
      q40_5 = "Election officials will discourage voting"),
    facet_label_wrap = 10,
    labels_size = 2.5,
    symmetric = F,
    add_totals = T
  )+
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Local Area", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.y.left = element_blank(), # remove y-axis text on 2nd plot
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_text(angle = 0, size = 8)
  )

# Instead of two separate plots, arrange into one plot while remaining separated
# use `ggpubr` package to print multiple plots side-by-side (or alt. row-by-row)
ggpubr::ggarrange(
  distrust.likert.az,
  distrust.likert.lcl,
  ncol = 2,
  common.legend = T,
  legend = "bottom", 
  font.label = list(family = "Palatino Linotype")
)

```

Similar to the trust scale items, the items on the distrust scale were associated with lower distrust in the treatment condition compared to the control. A quick glance at @fig-distrustlikert shows what appears to be a significant reduction in distrust, especially regarding the expectations that many votes wouldn't be counted, many voters would be turned away, and that election officials would discourage voters from casting a ballot. Interestingly, the expectation that there would be foreign interference in elections in Maricopa County, AZ were pretty resilient against the treatment. And again, responses for items pertaining to elections in one's local area are practically indistinguishable between treatment and control.

### Confidence in elections

Once trust and distrust were composed into unified variable of confidence in elections, I then examined distribution of the dependent variable followed by analysis of the effect of the treatment by comparison of mean differences for confidence in elections by experiment condition. Recall that on this scale, zero reflects insecurity (e.g., similar or equal values of trust and distrust), whereas values diverging from zero reflect confidence in either direction, trust or distrust respectively. Positive values reflect confidence as trust in elections administration (e.g., elections will be fair and accurate), whereas negative values reflect confidence as distrust in elections (i.e., expectation that election fraud will occur). To retain a meaningful zero, I rescaled the sum scores to range from -3 to 3[^9] and then superimposed two histograms of confidence scores: one derived from AZ items and another for scores derived from local area items. Doing this is illustrative of the relative distribution of confidence that differed depending on the location of elections referred to in the survey items (e.g., Maricopa County, AZ).

As revealed by review of the item responses, the distribution of confidence differed depending on whether survey items pertained to elections in Maricopa County, AZ, or elections within one's local area.

```{r}
#| label: fig-dist
#| fig-cap: "Overlapping histograms of Confidence in Elections for AZ and Local Area"
#| fig-pos: "H"

ggplot(df) +
  geom_histogram(
    aes(x = conf.az.rescaled , y = after_stat(density)),
    bins = 15,
    alpha = 0.2,
    color = 1,
    fill = "red"
  ) +
  geom_histogram(
    aes(x = conf.lcl.rescaled , y = after_stat(density)),
    bins = 15,
    alpha = 0.2,
    color = 1,
    fill = "blue"
  ) +
  geom_segment(
    x = round(mean(df$conf.az.rescaled), 2),
    y = 0,
    yend = 0.35,
    color = "firebrick",
    lwd = 1.4,
    linetype = 2
  ) +
  geom_segment(
    x = round(mean(df$conf.lcl.rescaled), 2),
    y = 0,
    yend = 0.35,
    color = "darkblue",
    lwd = 1.4,
    linetype = 2
  ) +
  annotate(
    "text",
    x = round(mean(df$conf.az.rescaled), 2),
    y = 0.40,
    label = paste("AZ =", round(mean(
      df$conf.az.rescaled
    ), 2), "(", round(sd(
      df$conf.az.rescaled
    ), 2), ")"),
    color = "firebrick"
  ) +
  annotate(
    "text",
    x = round(mean(df$conf.lcl.rescaled), 2),
    y = 0.43,
    label = paste("Local area =", round(mean(
      df$conf.lcl.rescaled
    ), 2), "(", round(sd(
      df$conf.lcl.rescaled
    ), 2), ")"),
    color = "darkblue"
  ) +
  labs(
    x = str_wrap("Confidence", width = 85),
    y = "Density",
    caption = str_wrap(
     "Mean (std.deviation). Difference = 0.20, CI [0.068, 0.33], t = 2.959, df=1283.733, p = 0.003",width = 95)) +
  theme_bw()

```

Overall, the sample expressed more trust in elections relative to distrust seeing as how the distribution of confidence is largely positive (@fig-dist). However, there's a clear difference between items that asked about elections in Maricopa county, AZ and those that asked about one's local area. Confidence in elections within one's local area positively differed from confidence in elections held in Maricopa County, AZ. A paired t-test suggests that this difference was statistically significant, though somewhat small (mean difference $= -0.33$, $95\%$ CI $[-0.38, -0.27]$, $t(1286) = -11.61$, $p < .001$; Cohen's $d = -0.32$, $95\%$ CI $[-0.38, -0.27]$). Nonetheless, respondents appeared less confident regarding elections in Maricopa County, AZ, but held more confidence in elections administration in their local area. This complements previous research that demonstrates a similar bias favorable to one's local area [@stewart2022].

Accordingly, when confidence levels are distinguished by experiment condition, the treatment vignette only had an effect upon survey items pertaining to elections in Maricopa County, AZ. Conducting an independent samples t-test (i.e., Welch Two Sample t-test) suggests that the effect of the treatment compared to the control condition is positive, statistically significant, yet substantially pretty small.

```{r}
#| label: tbl-ttest
#| tbl-cap: "Mean Difference Effect of Treatment Compared to Control Condition"
#| fig-pos: "h"
#| include: false

df |> 
  mutate(group = forcats::fct_relevel(group, "Treatment", after = 0L)) |> 
  rstatix::t_test(formula = conf.az.rescaled ~ group, var.equal = F, detailed = T) |>
  select(Difference = estimate, 
         conf.low, conf.high, 
         t = statistic, df, p, 
         -alternative, -method, -.y., -estimate1, -estimate2, -group1, -group2,
         -n1, -n2) |> 
  # insight::format_table(stars = T) |> 
  kableExtra::kbl(booktabs = T, digits = 4) |> 
  kableExtra::kable_styling(
    full_width = T,
    latex_options = c("hold_position"))

```

The results of @tbl-ttest show that, on a scale ranging from -3 to 3, the effect of the treatment is associated with a $0.20$ average difference in confidence in elections in Maricopa County, AZ, compared to average confidence among those in the control group. A simple difference of means test on this scale, however, does not lend itself well to interpretation; its hard to really grasp what a $0.20$ positive difference, on average, really amounts to. When the dependent variable is centered and scaled to have a mean = 0 and standard deviation = 1, standardized parameters permit interpretation of the treatment effect in terms of standard deviations.

> "tbl-mod1" here

The table above shows that effect of the treatment on confidence in elections is associated with a 0.16 standard deviation positive difference on average compared to confidence of those in the control group. I also include models where trust and distrust are regressed on the treatment condition separately in order to illustrate the divergent effect of the treatment. Nothing is more illustrative than graphs and lines, however.

> "fig-coef1" and "fig-coef2" here

Visualized in @fig-coef2, the treatment effect on both trust and distrust is clear. Although the effect is only there when questions pertained to Maricopa County, AZ. The treatment effect on trust and distrust is conditional on whether the survey questions inquired about elections in Maricopa County, AZ. Announcing efforts to recruit veterans and their families to work as election staff and volunteers increases confidence in elections administration in places outside of one's local area, but doesn't boost confidence in elections within one's local area. However, bear in mind that confidence in elections for one's local area is already higher than for those outside. Thus, there's already less insecurity concerning elections in one's local area, but publicized efforts to recruit veterans to work as election staff and volunteers is a small but positive step towards closing that gap in confidence for elections that occur elsewhere.

I now turn to results examining those who held onto the belief that the 2020 election results were illegitimate. I think check to see if recruiting veterans simply makes people feel safer at the polls.

\pagebreak

## Conclusion

### Limitations

As far as I'm aware, both the items, their wording, and their composition into measurement scales are unique to this study. This renders the composite scores for trust in elections sample dependent and limits comparability with other cross-sectional survey data.

-   limitations: interpretations of the treatment effect are limited.
    -   The sample may not be generalizable to the population.
    -   It's hard to say whether merely mentioning veterans is enough, as compared to explicitly naming veterans as the target of particular recruitment efforts.
-   Interpretation is also limited considering that no other particular group, or groups are compared directly against the veteran treatment vignette, i.e., additional vignettes for other comparable groups.
-   Moreover, results are limited by the survey questions, questionnaire design, and experimental stimulus (i.e., vignette) for two reasons. - First, the vignettes and many survey questions ask about Maricopa County, AZ specifically. Adding a specific county in the vignette adds in a factor that cannot be accounted for without additional treatments that eliminate the setting as a potential influence. Moreover, the specificity of the setting adds in even more unexplained error---some people may have attitudes about the county in question, others won't, while others may be muffed to consider a random county in the U.S. they've know nothing about. Adding the county undermines the confidence that treatment effects are solely attributable to veterans to an unknown degree. - Second, questions ask about one's own local community in addition to identical questions which asked about Maricopa County, AZ. Consequently, many questions within the questionnaire were duplicates with that one differences. This isn't uncommon in surveys, however in this case, it lengthened the survey to a degree that likely resulted in a higher drop off rate. More importantly, the quality of responses were likely diminished to some unknown extent. Although it is possible to compare questions asked about Maricopa County, AZ to questions about one's local area, the quality of that comparison is limited by the unknown extent of fatigue induced by answering the same questions twice. It wasn't just that some questions were asked multiple times, almost all of the questions were duplicated; after completing a long series on Maricopa County, participants then answered the same questions about the local community. The rate to which participants dropped from the survey is \_\_\_, which suggests fatigue as an important factor. Survey fatigue is a known issue that should be taken into consideration. Comparison of treatment effects between the MC series and the local area series is undermined by the unknown influence that fatigue would have on response choices. Any differences couldn't confidently be attributed to "my area" vs Maricopa county. We also can't determine whether treatment effects are sustained when comparing the Maricopa County series to the local area series asked subsequently. We would have to assume the setting as irrelevant, which we can't reasonably do.

\pagebreak

## Appendix A: Survey Experiment Vignettes and Survey Items {#sec-appendix-a}

::::: {#fig-vignettes .content-visible layout-ncol="2" when-format="html" unless-format="pdf"}
::: {#fig-treatment}
### Treatment Vignette

| **Local Military Veterans Recruited for Election Jobs in Maricopa County**
| 
|   PHOENIX (AP) — Election officials in Maricopa County, Arizona, announced a program designed to recruit military veterans and their family members from the community to serve as election administrators, including election polling place workers, temporary workers, and full-time staff. As the U.S. general elections in November near, election officials must fill several thousand temporary positions and hundreds of other open positions to ensure sufficient staffing for the 2024 elections and beyond.
| 
|   Army veteran Jordan Braxton just joined the elections workforce. Jordan believes their role is important to ensuring a secure, accurate, and transparent election, “Many places are short on staff this election cycle. I served my country in the Army, and I want to do my part as a veteran and a citizen to ensure that everyone trusts the process and the outcome of the election.”
:::

::: {#fig-control}
### Control Vignette

| **Local Residents Recruited for Election Jobs in Maricopa County**
| 
|   PHOENIX (AP) —Election officials in Maricopa County, Arizona, announced a program to recruit members of the community to serve as election administrators, including election polling place workers, temporary workers, and full-time staff. As the U.S. general elections in November near, election officials must fill several thousand temporary positions and hundreds of other open positions to ensure sufficient staffing for the 2024 elections and beyond.
| 
|   Jordan Braxton just joined the elections workforce. Jordan believes their role is important to ensuring a secure, accurate, and transparent election, “Many places are short on staff this election cycle. I want to do my part as a citizen to ensure that everyone trusts the process and the outcome of the election.”
:::

Fabricated Article Vignettes utilized as Experimental Stimulus
:::::

::::: {.content-hidden unless-format="pdf" layout="[0.45, -0.1, 0.45]"}
::: {#first-column}
### Treatment Vignette

| **Local Military Veterans Recruited for Election Jobs in Maricopa County**
| 
|   PHOENIX (AP) — Election officials in Maricopa County, Arizona, announced a program designed to recruit military veterans and their family members from the community to serve as election administrators, including election polling place workers, temporary workers, and full-time staff. As the U.S. general elections in November near, election officials must fill several thousand temporary positions and hundreds of other open positions to ensure sufficient staffing for the 2024 elections and beyond.
| 
|   Army veteran Jordan Braxton just joined the elections workforce. Jordan believes their role is important to ensuring a secure, accurate, and transparent election, “Many places are short on staff this election cycle. I served my country in the Army, and I want to do my part as a veteran and a citizen to ensure that everyone trusts the process and the outcome of the election.”
:::

::: {#second-column}
### Control Vignette

| **Local Residents Recruited for Election Jobs in Maricopa County**
| 
|   PHOENIX (AP) —Election officials in Maricopa County, Arizona, announced a program to recruit members of the community to serve as election administrators, including election polling place workers, temporary workers, and full-time staff. As the U.S. general elections in November near, election officials must fill several thousand temporary positions and hundreds of other open positions to ensure sufficient staffing for the 2024 elections and beyond.
| 
|   Jordan Braxton just joined the elections workforce. Jordan believes their role is important to ensuring a secure, accurate, and transparent election, “Many places are short on staff this election cycle. I want to do my part as a citizen to ensure that everyone trusts the process and the outcome of the election.”
:::
:::::

\pagebreak

## Appendix B: Sample Demographics {#sec-appendix-b}

The median age was 46 (mean age was 47). There was 51.7% (n = 658) women, 47% (n = 598) men, and approximately 1.3% who identified as either Non-binary/third gender (n = 7) or preferred not to say (n = 9). A large proportion of the sample identified as White or Caucasian (n = 975, 76.65%), while all other non-White respondents comprised 27.83% of the sample. Those who held a graduate level degree (e.g., Master's, Doctorate, or Professional level) comprised 13% of the sample; those with either degree at the Associate or Bachelor's level comprised 36.24%, while 22.25% had some college but no degree; and 28.46% had either a high school level or equivalent education or less than high school. The largest proportion of the sample identified as Democrat at 44.64%, followed by Republicans at 42.59%. The proportion of true Independents was 12.78%.

\pagebreak

\pagebreak

[^1]: I adopt a distinction made between *trustworthiness* and *trust* in elections [@stewart2022]. The "*worthiness*" of one's trust in the conduct and administration of elections is based on the extent that outcomes of an election reasonably follow the rules prescribed and can be adjudicated as such. To put it briefly, trustworthiness is built by the structure, procedures, and practices of the institution, in this case election administration. The public's trust in elections, however, is amendable to an indefinite number of factors that may be unrelated to the formal structure or procedures of election administration. To illustrate, I can recognize that my car is trustworthy prior to ever driving it because it is structurally sound; it passes whatever criteria upon inspection, it is sufficiently fueled, and appears to be in generally good working order. I have every reason to deem it *worthy of my trust*. However, I don't trust my car because I am sure it is haunted.

[^2]: There's little that could justify a conceptual distinction between public trust pre-election and public trust post-election. The temporal element renders the difference between pre- and post-election more operative than conceptual at a foundational level.

[^3]: Note that election insecurity is used in place of electoral confidence simply to emphasize the anticipatory nature of the attitudes in question. Election insecurity is taken to denote a lack of confidence.

[^4]: Demographic breakdown of the sample are included in Appendix B.

[^5]: Complete text of treatment and control vignettes are included in the Appendix

[^6]: Note that I refer to the items as either 'AZ items' or 'local items' in order to distinguish the location in which they pertain.

[^7]: Survey item wording and responses are included in Appendix A.

[^8]: The methodological concerns between the use of (unit weighted) sum or mean scores compared to the use of estimated factor scores for Likert scales composed of multiple Likert items are legion. The use of either computed sum scores or estimated factor scores as a general practice is continuously debated [@mcneish2020; @widaman2022; @mcneish2023; @widaman2024], yet there is no general consensus on the most appropriate method generally. The use of either appears to largely depend on the researcher's particular research objective, theoretical suppositions, and necessity. Another common concern is the inappropriate treatment of ordinal variables as interval which levy strong assumptions, e.g., each item contributes equally to the variable being measured and that ordinal response categories are equidistant as though on an interval scale [@wang2013; for contrast, see @robitzsch2020].

[^9]: This range is mostly arbitrary, as a range from -1 to 1 works much the same. When relying on simple sum total score, the scale ranges from -15 to 15. So a single point increase from 0 to 1 may reflect a combination of a single "Very confident" response on a trust item compensated by a single "Somewhat likely" response on a distrust item (i.e., 3-2 = 1), or some other equivalent combination. Although the use of mean scores would place scores back onto the response scale metric (e.g., from 0 to 3, reflective of "not at all" to "very"), this results in a unipolar scale (from 0 to 3). The bipolar scale resulting from the composition of positive trust and negative distrust engenders meaning to zero and negative values. As such, composite mean scores become inappropriate and interpretation of scores would no longer be feasible. In other words, negative values on the scale for confidence in elections hold substantive meaning (e.g., distrust, absent trust), which makes transforming scores to fit a unipolar scale inappropriate.