### 2025-03-13 notebooks/figures.qmd

---
title: "Figures"
execute:
  enabled: true
---

```{r}
#| label: packages
#| message: false

set.seed(1234)
library(tidyverse)
library(here)
library(patchwork)
library(pewmethods)
library(ggstats)

```

```{r}
#| label: data-import

# Load data
load(here::here('data', "research-paper-data-20241029.Rdata"))

# shorten and 'unify' factor levels
# Note: this removes variable labels
data <- data |>
  # use `pewmethods` to preserve order of factor levels
  mutate(across(c(q19, q20, q21, q22, q23, q24,
                  q30, q31, q32, q33, q34, q35),
                ~pewmethods::fct_case_when(
    .x %in% c("Not at all confident",
              "Not at all committed",
              "Not at all concerned") ~ "Not at all confident",
    .x %in% c("Not too confident",
              "Not too committed",
              "Not too concerned",
              "Somewhat unconcerned") ~ "Not too confident",
    .x %in% c("Somewhat confident",
              "Somewhat committed",
              "Somewhat concerned") ~ "Somewhat confident",
    .x %in% c("Very confident",
              "Very committed",
              "Very concerned") ~ "Very confident",
    TRUE ~ NA
  ))) |> 
  # unify factor levels
  mutate(across(c(q25, q36), ~ pewmethods::fct_case_when(
    .x == "Very concerned" ~ "Very concerned",
    .x == "Somewhat concerned" ~ "Somewhat concerned",
    .x %in% c("Not too concerned", "Somewhat unconcerned") ~ "Not too concerned",
    .x == "Not at all concerned" ~ "Not at all concerned",
    TRUE ~ NA))) |>
  # reverse factor levels
  mutate(across(c(q25, q36), ~forcats::fct_rev(.)))

```

```{r}
#| label: setup-theme

# set ggplot theme that will be used for every plot
# theme_get()
old <- ggplot2::theme_set(ggplot2::theme_bw())  # capture current theme

# set/update new theme
ggplot2::theme_set(
  new = ggplot2::theme_bw(
    base_family = 'TeX Gyre Pagella')+ # set font of ggplots
    theme(legend.position = 'bottom')  # set legend position for all plots
)

```

```{r}
#| label: custom-function

# custom function used to create Likert Plots

# To avoid excessive copy/paste, I made a function to create Likert plots
# it is mostly a wrapper around `ggstats::gglikert`, but with a specific set up
# particular colors, theme, etc. 
likert_plot <- function(data, x, ...,  
                        symmetric = FALSE,
                        variable_labels = NULL,
                        label_size = 2.5,
                        vline = FALSE,
                        title = waiver(), 
                        subtitle = waiver(),
                        caption = waiver(),
                        xlab = waiver()){

  p <- ggstats::gglikert(
    data = data, include = {{ x }}, ...,
    variable_labels = variable_labels,
    symmetric = symmetric)+

    # customize color
    ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
    # custom theme
    theme_bw(base_family = "TeX Gyre Pagella")+
    theme(
      legend.position = 'bottom',  # place legend on bottom
      axis.title.y.left = element_blank(), # blank out y-axis label
      axis.text.x = element_blank(), # remove percentage text along x-axis
      strip.text.y.right = element_text(angle = 0) # make facet label horizontal
        )

    if (vline==TRUE) {
    p <- p + geom_vline(xintercept = 0, color = 'black', linewidth = 1.2)
  } else {
      p
  }

  p <- p + ggplot2::labs(..., title = title, subtitle=subtitle, caption=caption, x = xlab)
  p

}

```

```{r}
#| eval: true
#| results: hide

# assign items to particular vectors to avoid copy/pasting
trust.az.items <- data |> select(q19, q20, q21, q22, q24) |> colnames() |> dput()
trust.lcl.items <- data |> select(q30, q31, q32, q33, q35)|> colnames() |> dput()
distrust.az.items <- data |> select(q28_1:q28_5)|> colnames() |> dput()
distrust.lcl.items <- data |> select(q40_1:q40_5)|> colnames() |> dput()

# variable labels for az and local area items, respectively
trust.az.varlabels <- c(
  q19 = "Votes counted as intended",
  q20 = "Election officials will do good job",
  q21 = "Election workers will be committed",
  q22 = "Voting process will be fair",
  q24 = "Election technology will be secure"
)

distrust.az.varlabels <- c(
  q28_1 = "Voter fraud will occur",
  q28_2 = "Many votes won't be counted",
  q28_3 = "Many voters will be turned away",
  q28_4 = "Foreign Interference",
  q28_5 = "Election officials will discourage voting"
)

trust.lcl.varlabels <- c(
  q30 = "Votes counted as intended",
  q31 = "Election officials will do good job",
  q32 = "Election workers will be committed",
  q33 = "Voting process will be fair",
  q35 = "Election technology will be secure"
)

distrust.lcl.varlabels <- c(
  q40_1 = "Voter fraud will occur",
  q40_2 = "Many Votes won't be counted",
  q40_3 = "Many voters will be turned away",
  q40_4 = "Foreign Interference",
  q40_5 = "Election officials will discourage voting"
)

```

## Trust in Elections across sample

```{r}
#| label: fig-p1
#| fig-cap: "Trust in Maricopa County, AZ, Elections"

# plot of trust in elections

# plot of trust in AZ elections 
trust.likertplot.az <- data |> likert_plot(
  x = all_of(trust.az.items),
  variable_labels = trust.az.varlabels,
  title = "Trust in Maricopa County, AZ, Elections"
)

trust.likertplot.az

```

```{r}
#| label: fig-p2
#| fig-cap: "Trust in Local Area Elections"

# plot of trust in Local Area elections 
trust.likertplot.lcl <- likert_plot(
  data = data, 
  x = all_of(trust.lcl.items),
  variable_labels = trust.lcl.varlabels,
  title = "Trust in Local Area Elections")

trust.likertplot.lcl

```

```{r}
#| label: fig-patch1
#| fig-cap: "Trust in Elections"

# Instead of two separate plots, arrange into one plot while remaining separated
p1.trust <- trust.likertplot.az / trust.likertplot.lcl+
  patchwork::plot_annotation(caption = "Missing values omitted")+
  patchwork::plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

p1.trust

```

## Distrust in elections across sample

```{r}
#| label: fig-p3
#| fig-cap: "Distrust in Maricopa County, AZ Elections"

# responses to distrust in AZ election items, overall sample
distrust.likert.az <- likert_plot(
  data, 
  x = all_of(distrust.az.items), 
  variable_labels = distrust.az.varlabels,
  title = "Distrust in Maricopa County, AZ Elections", 
  subtitle = str_wrap(
    "How likely do you think any or all of the following will happen during this year´s elections in Maricopa County, AZ?",width = 95))

distrust.likert.az

```

```{r}
#| label: fig-p4
#| fig-cap: "Distrust in Local Area Elections"

# responses to distrust in local election items, overall sample
distrust.likert.lcl <- likert_plot(
  data,
  x = all_of(distrust.lcl.items),
  variable_labels = distrust.lcl.varlabels,
  title = "Distrust in Local Area Elections",
  subtitle = str_wrap(
    "How likely do you think any or all of the following will happen during this year´s elections in your local area?",
    width = 95
  )
)

distrust.likert.lcl

```

```{r}
#| label: fig-patch2
#| fig-cap: "Distrust in Elections"

p2.distrust <- distrust.likert.az/distrust.likert.lcl+
  patchwork::plot_annotation(caption = "Missing values omitted")+
  patchwork::plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

p2.distrust

```

## Concerns for Violence and Confidence in Voter Safety Across Sample

```{r}
#| label: fig-p5

# plot of safety concerns and confidence ::::::::::::::::::::::::::::

safety.concerns.likert <- likert_plot(
  data, x = c(q25, q36), 
  y_label_wrap = 15, 
  variable_labels = c(
    q25 = 'Thinking about Maricopa County, AZ',
    q36 = "Thinking about Local Area"), 
  title = "Concern for Voter Intimidation, Threats, or Violence",
  subtitle = str_wrap(
    "How concerned should voters feel about potential violence, threats of violence, or intimidation while voting in person at their local polling place?", width = 99))

safety.concerns.likert

```

```{r}
#| label: fig-p6

# q26 In-person voter safety at polling sites
safety.conf.likert <- likert_plot(
  data,
  x = c(q26, q37),
  y_label_wrap = 15,
  variable_labels = c(q26 = "In Maricopa County, AZ", q37 = "In your local area"),
  title = str_wrap("Confidence for In-person voter safety at Polling Sites", width = 95),
  subtitle = str_wrap(
    "How confident, if at all, are you that in person polling places will be safe places for voters to cast their ballots during the upcoming elections in November?",
    width = 95
  )
)

safety.conf.likert

```

```{r}
#| label: fig-patch3

p3.safety <- safety.concerns.likert/safety.conf.likert+
  patchwork::plot_annotation(caption = "Missing values omitted")+
  patchwork::plot_layout(guides = "keep") &
  theme(legend.position = "bottom")

```

## Trust and Distrust by Education, Gender, and Partisanship

```{r}
#| label: fig-p7

# plots of trust in AZ elections by educational attainment :::::::::: 
trust.az.educ.likert <- data |>   
  filter(!is.na(educ_4cat)) |> 
  mutate(
    educ_3cat = pewmethods::fct_case_when(
      educ_4cat %in% c("H.S. or less") ~ "H.S. or less",
      educ_4cat %in% c("Some college no degree") ~ "Some college no Degree",
      educ_4cat %in% c("Postgraduate degree", "College degree") ~ "College Graduate")
    ) |>  
  likert_plot(x = all_of(trust.az.items), y = 'educ_3cat', facet_rows = vars(.question),
              facet_label_wrap = 15,
              variable_labels = trust.az.varlabels,
              subtitle = "Maricopa County, AZ")
```

```{r}
#| label: fig-p8

# plots of trust in Local area elections by educational attainment :: 
trust.lcl.educ.likert <- data |>  
  filter(!is.na(educ_4cat)) |>  
  mutate(
    educ_3cat = pewmethods::fct_case_when(
      educ_4cat %in% c("H.S. or less") ~ "H.S. or less",
      educ_4cat %in% c("Some college no degree") ~ "Some college no Degree",
      educ_4cat %in% c("Postgraduate degree", "College degree") ~ "College Graduate")
    ) |>  
  likert_plot(
    x = all_of(trust.lcl.items), y = 'educ_3cat', facet_rows = vars(.question),
    facet_label_wrap = 15, variable_labels = trust.lcl.varlabels,
    subtitle = "Local Area"
  )
```

```{r}
#| label: fig-patch4

p4.trust.az.educ <- trust.az.educ.likert + trust.lcl.educ.likert + 
    patchwork::plot_layout(guides = "collect") &
    theme(legend.position = "bottom")

```

```{r}
#| label: fig-p9

# plot of distrust in AZ elections by educational attainment :::::::: 
distrust.az.educ.likert <- data |> 
  filter(!is.na(educ_4cat)) |> 
  mutate(
    educ_3cat = pewmethods::fct_case_when(
      educ_4cat %in% c("H.S. or less") ~ "H.S. or less",
      educ_4cat %in% c("Some college no degree") ~ "Some college no Degree",
      educ_4cat %in% c("Postgraduate degree", "College degree") ~ "College Graduate")
    ) |>  
  likert_plot(x = all_of(distrust.az.items), y = 'educ_3cat', facet_rows=vars(.question),
              facet_label_wrap = 15, variable_labels = distrust.az.varlabels,
              subtitle = "Maricopa County, AZ")
```

```{r}
#| label: fig-p10

# plot of distrust items by educational attainment
distrust.lcl.educ.likert <- data |> 
  filter(!is.na(educ_4cat)) |> 
  mutate(
    educ_3cat = pewmethods::fct_case_when(
      educ_4cat %in% c("H.S. or less") ~ "H.S. or less",
      educ_4cat %in% c("Some college no degree") ~ "Some college no Degree",
      educ_4cat %in% c("Postgraduate degree", "College degree") ~ "College Graduate")
    ) |>   
  likert_plot(
    x = all_of(distrust.lcl.items), y = 'educ_3cat', facet_rows = vars(.question),
    facet_label_wrap = 15, variable_labels = distrust.lcl.varlabels,
    subtitle = "Local Area"
  )

```

```{r}
#| label: fig-patch5

p5.distrust.az.educ <- distrust.az.educ.likert + distrust.lcl.educ.likert +
    patchwork::plot_layout(guides = "collect") & 
    theme(legend.position = "bottom")

p5.distrust.az.educ

```

```{r}
#| label: fig-p11

# plots of trust and distrust in AZ elections by gender ::::::::::::: 
trust.az.gender.likert <- data |>  
  filter(!is.na(gender_3cat), gender_3cat != 'Other/Refused') |> 
  likert_plot(
    x = all_of(trust.az.items), y = 'gender_3cat', facet_rows = vars(.question),
    facet_label_wrap = 15, variable_labels = trust.az.varlabels,
    subtitle = "Maricopa County, AZ Elections"
  )
```

```{r}
#| label: fig-p12

# plots of distrust in AZ elections by gender ::::::::::::: 
distrust.az.gender.likert <- data |> 
  filter(!is.na(gender_3cat), gender_3cat != 'Other/Refused') |>
  likert_plot(
    x = all_of(distrust.az.items), y = 'gender_3cat', facet_rows = vars(.question),
    facet_label_wrap = 15, variable_labels = distrust.az.varlabels,
    subtitle = "Maricopa County, AZ Elections"
  )
```

```{r}
#| label: fig-patch6

p6.trust.az.gender <- trust.az.gender.likert / distrust.az.gender.likert +
  patchwork::plot_annotation(
    caption = str_wrap(
      "Gender Non-binary (n=7) and those who preferred not to say (n=9) response categories omitted for lack of observations. Missing values omitted (n=15)",
      width = 150
    ),
    theme = theme(text = element_text(family = "Palatino Linotype"))
  ) +
  patchwork::plot_layout(guides = 'keep') &
  theme(legend.position = "bottom")

p6.trust.az.gender

```

```{r}
#| label: fig-p13

# plot of trust in local area elections by gender
trust.lcl.gender.likert <- data |>  
  filter(!is.na(gender_3cat), gender_3cat != 'Other/Refused') |> 
  likert_plot(
    x = all_of(trust.lcl.items), y = 'gender_3cat',
    facet_rows = vars(.question),
    facet_label_wrap = 15, variable_labels = trust.lcl.varlabels,
    subtitle = "Local Area Elections"
  )
```

```{r}
#| label: fig-p14

# plot of distrust in local area elections by gender
distrust.lcl.gender.likert <- data |> 
  filter(!is.na(gender_3cat), gender_3cat != 'Other/Refused') |> 
  likert_plot(
    x = all_of(distrust.lcl.items), y = 'gender_3cat',
    facet_rows = vars(.question),
    facet_label_wrap = 15, variable_labels = distrust.lcl.varlabels,
    subtitle = "Local Area Elections"
  )
```

```{r}
#| label: fig-patch7

p7.trust.lcl.gender <- trust.lcl.gender.likert/distrust.lcl.gender.likert+
  patchwork::plot_annotation(
    caption = str_wrap("Gender Non-binary (n=7) and those who preferred not to say (n=9) response categories omitted for lack of observations. Missing values omitted (n=15)",
                       width = 150), 
    theme = theme(text = element_text(family = "Palatino Linotype")))+
    patchwork::plot_layout(guides = 'keep') & 
    theme(legend.position = "bottom")
```

```{r}
#| label: fig-p15

# plot of trust in AZ elections by partisanship ::::::: 
trust.az.party.likert <- data |>  
  filter(!is.na(partyid_3cat)) |> 
  likert_plot(
    x = all_of(trust.az.items), y = 'partyid_3cat',
    facet_rows = vars(.question), 
    facet_label_wrap = 15,
    variable_labels = trust.az.varlabels,
    subtitle = "Maricopa County, AZ Elections"
  )
```

```{r}
#| label: fig-p16

# plot of distrust in AZ elections by partisanship ::::::: 
distrust.az.party.likert <- data |> 
  filter(!is.na(partyid_3cat)) |> 
  likert_plot(
    x = all_of(distrust.az.items), y = 'partyid_3cat',
    facet_rows = vars(.question), 
    facet_label_wrap = 15,
    variable_labels = distrust.az.varlabels,
    subtitle = "Maricopa County, AZ Elections"
  )
```

```{r}
#| label: fig-patch8

p8.trust.az.party <- trust.az.party.likert/distrust.az.party.likert+
    patchwork::plot_layout(guides = 'keep') & 
    theme(legend.position = "bottom")

p8.trust.az.party

```

```{r}
#| label: fig-p17

# plot of trust and distrust in local area elections by partisanship  
trust.lcl.party.likert <- data |>  
  filter(!is.na(partyid_3cat)) |> 
  likert_plot(
    x = all_of(trust.lcl.items), y = 'partyid_3cat',
    facet_rows = vars(.question), 
    facet_label_wrap = 15,
    variable_labels = trust.lcl.varlabels,
    subtitle = "Local Area Elections"
  )
```

```{r}
#| label: fig-p18

# plot of distrust in local area elections by partisanship  
distrust.lcl.party.likert <- data |> 
  filter(!is.na(partyid_3cat)) |>  
  likert_plot(
    x = all_of(distrust.lcl.items), y = 'partyid_3cat',
    facet_rows = vars(.question), 
    facet_label_wrap = 15,
    variable_labels = distrust.lcl.varlabels,
    subtitle = "Local Area Elections"
  )
```

```{r}
#| label: fig-patch9

p9.trust.lcl.party <- trust.lcl.party.likert/distrust.lcl.party.likert+
    patchwork::plot_layout(guides = 'keep') & 
    theme(legend.position = "bottom")

p9.trust.lcl.party

```

## Trust and Distrust by Treatment Condition

```{r}
#| label: fig-p19

# plot of trust in AZ elections by treatment condition
trust.likert.treat.az <- data |>
  likert_plot(x = all_of(trust.az.items), y = 'group', 
    facet_rows = vars(.question),
    subtitle = "Maricopa County, Arizona")+
  theme(strip.text.y.right = element_blank())

trust.likert.treat.az

```

```{r}
#| label: fig-p20

# plot of trust in local elections by treatment condition
trust.likert.treat.lcl <- data |>
  likert_plot(x = all_of(trust.lcl.items), y = 'group', 
    facet_rows = vars(.question),
    facet_label_wrap = 10,
    variable_labels = trust.lcl.varlabels, 
    subtitle = "Local Area")+
  theme(strip.text.y.right = element_blank()) +
  theme(
    axis.text.y.left = element_blank(), # remove y-axis text on 2nd plot
    strip.text.y.right = element_text(angle = 0, size = 8)
  )

trust.likert.treat.lcl

```

```{r}
#| label: fig-patch10
#| fig-cap: "Trust in Election Administration by Experiment Condition"
#| fig-height: 4.5
#| fig-pos: "h"

p10.trust.treat <- trust.likert.treat.az + trust.likert.treat.lcl+
  patchwork::plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

p10.trust.treat

```

```{r}
#| label: fig-p21

# distrust in AZ elections by experimental condition
distrust.az.treat.likert <- data |>
  likert_plot(x = all_of(distrust.az.items), y = 'group', 
    facet_rows = vars(.question),
    subtitle = "Maricopa County, AZ")+
  theme(strip.text.y.right = element_blank())
```

```{r}
#| label: fig-p22

# distrust in local area elections by treatment condition
distrust.lcl.treat.likert <- data |>
  likert_plot(x = all_of(distrust.lcl.items), y = 'group', 
    facet_rows = vars(.question),
    facet_label_wrap = 10,
    variable_labels = distrust.lcl.varlabels,
    subtitle = "Local Area")+
  theme(strip.text.y.right = element_blank()) +
  theme(
    axis.text.y.left = element_blank(), # remove y-axis text on 2nd plot
    strip.text.y.right = element_text(angle = 0, size = 8)
  )
```

```{r}
#| label: fig-patch11
#| fig-cap: "Disrust in Election Administration by Experiment Condition"
#| fig-height: 4.5
#| fig-width: 8
#| fig-pos: "H"

p11.distrust.treat <- distrust.az.treat.likert + distrust.lcl.treat.likert+
  patchwork::plot_layout(guides = 'collect')

p11.distrust.treat

```

## trust and distrust by partisanship and treatment

```{r}
#| label: fig-p23

# plots of trust and distrust by partisanship and treatment :::::::::

# plot of trust in AZ elections by experiment condition and party id
trust.az.treat.partyid <- data |> 
  filter(!is.na(partyid_3cat)) |>
  likert_plot(
    x = all_of(trust.az.items),
    y = 'partyid_3cat',
    facet_rows = vars(.question),
    facet_cols = vars(group),
    facet_label_wrap = 15,
    variable_labels = trust.az.varlabels,
    title = "Trust in Maricopa County, AZ Elections by Treatment and Partisanship")
```

```{r}
#| label: fig-p24

# plot of trust in local elections by experiment condition and party id
trust.lcl.treat.partyid <- data |> 
  filter(!is.na(partyid_3cat)) |>
  likert_plot(
    x = all_of(trust.lcl.items),
    y = 'partyid_3cat',
    facet_rows = vars(.question),
    facet_cols = vars(group),
    facet_label_wrap = 15,
    variable_labels = trust.lcl.varlabels,
    title = "Trust in Local Area Elections by Treatment and Partisanship")
```

```{r}
#| label: fig-patch12

p12.trust.treat.partyid <- trust.az.treat.partyid / trust.lcl.treat.partyid+
  patchwork::plot_layout(guides = 'collect')

p12.trust.treat.partyid

```

## Trust and Distrust by Legitimacy beliefs and Treatment Condition

```{r}
#| label: fig-p25

# plots of trust and distrust by legitimacy and treatment :::::::::::

# trust in AZ elections by experiment conditions among those who believe the 2020
# election results were illegitimate
trust.az.treat.legit <- data |> 
  filter(!is.na(partyid_3cat), !is.na(q7)) |>
  filter(q7 == "Not legitimate") |> 
  likert_plot(
    x = all_of(trust.az.items), 
    y = 'group',
    facet_rows = vars(.question),
    facet_cols = vars(q7),
    facet_label_wrap = 15,
    variable_labels = trust.az.varlabels,
    subtitle = str_wrap("Maricopa County, AZ",
      width = 95) 
  )+
  theme(text = element_text(family = 'TeX Gyre Pagella'),
        strip.text.y.right = element_blank())
```

```{r}
#| label: fig-p26

# trust in local elections by experiment conditions among those who believe the
# 2020 election results were illegitimate
trust.lcl.treat.legit <- data |> 
  filter(!is.na(partyid_3cat), !is.na(q7)) |>
  filter(q7 == "Not legitimate") |> 
  likert_plot(
    x = all_of(trust.lcl.items), 
    y = 'group',
    facet_rows = vars(.question),
    facet_cols = vars(q7),
    facet_label_wrap = 15,
    variable_labels = trust.lcl.varlabels,
    subtitle = str_wrap("Local Area",
      width = 95) 
  )+
  theme(
    text = element_text(family = 'TeX Gyre Pagella'),
    axis.text.y.left = element_blank())
```

```{r}
#| label: fig-patch13
#| fig-cap: "Trust in elections by treatment among those who refute the 2020 election results"
#| fig-height: 4.5
#| fig-width: 8

p13.trust.treat.legit <- trust.az.treat.legit + trust.lcl.treat.legit+
  patchwork::plot_layout(guides = 'collect')

p13.trust.treat.legit

```

```{r}
#| label: fig-p27

# distrust in AZ elections by experiment conditions among those who believe the
# 2020 election results were illegitimate
distrust.az.treat.legit <- data |> 
  filter(!is.na(partyid_3cat), !is.na(q7)) |>
  filter(q7 == "Not legitimate") |> 
  likert_plot(
    x = all_of(distrust.az.items), 
    y = 'group',
    facet_rows = vars(.question),
    facet_cols = vars(q7),
    facet_label_wrap = 15,
    variable_labels = distrust.az.varlabels,
    subtitle = str_wrap("Maricopa County, AZ",
      width = 95) 
  )+
  theme(
    text = element_text(family = 'TeX Gyre Pagella'),
    strip.text.y.right = element_blank())

```

```{r}
#| label: fig-p28

# distrust in local elections by experiment conditions among those who believe
# the 2020 election results were illegitimate
distrust.lcl.treat.legit <- data |> 
  filter(!is.na(partyid_3cat), !is.na(q7)) |>
  filter(q7 == "Not legitimate") |> 
  likert_plot(
    x = all_of(distrust.lcl.items), 
    y = 'group',
    facet_rows = vars(.question),
    facet_cols = vars(q7),
    facet_label_wrap = 15,
    variable_labels = distrust.lcl.varlabels,
    subtitle = str_wrap("Local Area",
      width = 95) 
  )+
  theme(
    text = element_text(family = 'TeX Gyre Pagella'),
    axis.text.y.left = element_blank())

```

```{r}
#| label: fig-patch14
#| fig-cap: "Distrust in elections by treatment among those who refute the 2020 election results"
#| fig-height: 4.5
#| fig-width: 8

p14.distrust.treat.legit <- distrust.az.treat.legit + distrust.lcl.treat.legit+
  patchwork::plot_layout(guides = 'collect')

p14.distrust.treat.legit

```

## Concerns for Violence and Confidence in Voter Safety by Treatment

```{r}
#| label: fig-p29

# plot for safety concerns by treatment :::::::::::::::::::::::::::::

# Because q25 measures levels of concern/worry, it cannot be included in the
# same Likert plot as q26 as that measures reported confidence that voters will
# be safe to vote in-person

# safety concerns by treatment
safety.concern.treat.likert <- data |>
  # Need to slightly modify text of factor levels
  mutate(across(
    c(q25, q36),
    ~ pewmethods::fct_case_when(
      .x == "Very concerned" ~ "Very concerned",
      .x == "Somewhat concerned" ~ "Somewhat concerned",
      .x %in% c("Not too concerned", "Somewhat unconcerned") ~ "Not too concerned",
      .x == "Not at all concerned" ~ "Not at all concerned",
      TRUE ~ NA
    )
  )) |>
  likert_plot(
    x = c(q25, q36),
    y = 'group',
    facet_rows = vars(.question),
    facet_label_wrap = 15,
    variable_labels = c(q25 = 'Thinking about Maricopa County, AZ', q36 = "Thinking about Local Area"),
    title = "Concern for Voter Intimidation, Threats, or Violence by Treatment Condition",
    subtitle = str_wrap(
      "How concerned should voters feel about potential violence, threats of violence, or intimidation while voting in person at their local polling place?",
      width = 99
    )
  )
```

```{r}
#| label: fig-p30

# q26 In-person voter safety at election sites by treatment
safety.conf.treat.likert <- data |>
  likert_plot(
    x = c(q26, q37), 
    y = 'group',
    facet_label_wrap = 15,
    facet_rows = vars(.question),
    variable_labels = c(
      q26 = "In Maricopa County, AZ",
      q37 = "In your local area"
    ),
    title = str_wrap("Confidence for In-person voter safety at Polling Sites by Treatment", width = 95),
    subtitle = str_wrap("How confident, if at all, are you that in person polling places will be safe places for voters to cast their ballots during the upcoming elections in November?", width = 95)
  )

```

```{r}
#| label: fig-patch15

p15.safety.treat <- safety.concern.treat.likert/safety.conf.treat.likert+
  patchwork::plot_layout(guides = "keep")

```

## Distribution of Confidence in Elections

```{r}
#| label: fig-dist
#| fig-cap: "Overlapping Distribution of Confidence in Elections for AZ and Local Area"
#| fig-pos: "H"

ggplot(data) +
  geom_histogram(
    aes(x = conf.az.rescaled , y = after_stat(density)),
    bins = 15,
    alpha = 0.2,
    color = 1,
    fill = "red"
  ) +
  geom_histogram(
    aes(x = conf.lcl.rescaled , y = after_stat(density)),
    bins = 15,
    alpha = 0.2,
    color = 1,
    fill = "blue"
  ) +
  geom_segment(
    x = round(mean(data$conf.az.rescaled), 2),
    y = 0,
    yend = 0.35,
    color = "firebrick",
    lwd = 1.4,
    linetype = 2
  ) +
  geom_segment(
    x = round(mean(data$conf.lcl.rescaled), 2),
    y = 0,
    yend = 0.35,
    color = "darkblue",
    lwd = 1.4,
    linetype = 2
  ) +
  annotate(
    "text",
    x = round(mean(data$conf.az.rescaled), 2),
    y = 0.40,
    label = paste("AZ =", round(mean(
      data$conf.az.rescaled
    ), 2), "(", round(sd(
      data$conf.az.rescaled
    ), 2), ")"),
    color = "firebrick"
  ) +
  annotate(
    "text",
    x = round(mean(data$conf.lcl.rescaled), 2),
    y = 0.43,
    label = paste("Local area =", round(mean(
      data$conf.lcl.rescaled
    ), 2), "(", round(sd(
      data$conf.lcl.rescaled
    ), 2), ")"),
    color = "darkblue"
  ) +
  labs(
    x = str_wrap("Confidence", width = 85),
    y = "Density",
    caption = str_wrap(
     "Mean (std.dev). Mean difference = 0.33, CI [0.27, 0.38], t = 11.613, data = 1286, p < 0.01",
     width = 95)) +
  theme_bw(base_family = "TeX Gyre Pagella")

```

## Coefficient Plot of Treatment on Confidence by Location

```{r}
#| label: fig-coef1
#| layout-nrow: 2
#| fig-height: 3.5
#| fig-pos: 'H'
#| fig-cap: "Confidence in Elections by Experiment Condition"
#| fig-subcap: 
#|   - "Confidence by Treatment Condition"
#|   - "Trust and Distrust by Treatment Condition"

# coef plot of treatment effect on confidence distinguished by location
conf.coef1 <- jtools::plot_summs(
  lm(conf.az.rescaled ~ group, data = data), 
  lm(conf.lcl.rescaled ~ group, data = data), 
  center = T, scale = T, transform.response = T,
  model.names = c("Conf.AZ", "Conf.Local"),
  coefs = c("Treatment" = "groupTreatment"), 
  colors = c("red", "black"),
  point.shape = c(8, 1))+
  ggplot2::expand_limits(x = c(-0.25, 0.25))+
  labs(y = "")+
  theme_bw(base_family = "TeX Gyre Pagella")

# Coef plots of trust and distrust models for AZ and Local items
conf.coef2 <- jtools::plot_summs(
  lm(trust.scores.az ~ group, data = data),
  lm(distrust.scores.az ~ group, data = data),
  lm(trust.scores.lcl ~ group, data = data),
  lm(distrust.scores.lcl ~ group, data = data),
  center = T, scale = T, transform.response = T,
  model.names = c("Trust.AZ", "Distrust.AZ", "Trust.Local", "Distrust.Local"),
  coefs = c("Treatment" = "groupTreatment"),
  colors = c("red", "red", "black", "black"),
  point.shape = c(8, 8, 1, 1)
)+
  ggplot2::expand_limits(x = c(-0.25, 0.25))+
  labs(y = "")+
  theme_bw(base_family = "TeX Gyre Pagella")

conf.coef1
conf.coef2

```

## Polychoric Item Correlation Plot

```{r}
#| label: fig-polycor
#| fig-cap: "Polychoric Item Correlation Matrix"
#| fig-pos: 'h'

# Get polychoric item correlation matrix.
item.cor <- data |>
  select(q19:q24, -q23, q28_1:q28_5, q30:q35, -q34, q40_1:q40_5) |> 
  # ensure factors are ordered factors
  mutate(across(where(is.factor), ~factor(., ordered = T))) |> 
  # adjust variable column names of items
  dplyr::rename(
    trustvote_az     = q19,
    trustadmin_az    = q20,
    truststaff_az    = q21,
    trustprcss_az    = q22,
    trusttech_az     = q24,
    votefraud_az     = q28_1,
    miscount_az      = q28_2,
    turnedaway_az    = q28_3,
    foreigntamp_az   = q28_4,
    discourage_az    = q28_5,
    trustvote_lcl    = q30,
    trustadmin_lcl   = q31,
    truststaff_lcl   = q32,
    trustprcss_lcl   = q33,
    trusttech_lcl    = q35,
    votefraud_lcl    = q40_1,
    miscount_lcl     = q40_2,
    turnedaway_lcl   = q40_3,
    foreigntamp_lcl  = q40_4,
    discourage_lcl   = q40_5
  ) |> 
  correlation::correlation(
    select = 
      c("trustvote_az", "trustadmin_az", "truststaff_az", "trustprcss_az", "trusttech_az",
        "votefraud_az", "miscount_az", "turnedaway_az", "foreigntamp_az", "discourage_az"),
    select2 =
      c("trustvote_lcl", "trustadmin_lcl", "truststaff_lcl", "trustprcss_lcl","trusttech_lcl",
        "votefraud_lcl", "miscount_lcl", "turnedaway_lcl", "foreigntamp_lcl","discourage_lcl"),
    method = "polychoric") |>
  summary()

# plot item correlations
plot(item.cor) + theme(axis.text.x = element_text(angle = 45, vjust = 0.6))

```