### 2025-07-04 supplemental/appendix_D.qmd

---
title: "Appendix D"
subtitle: "Predicted Probabilities of Safety Concerns by Treatment Condition"
format:
  typst: 
    toc: false
    keep-typ: true
    papersize: us-letter
    mainfont: TeX Gyre Pagella
    fontsize: 11pt
execute: 
  echo: false
---

```{r}
#| label: setup
#| include: false
#| echo: false

set.seed(1234)

# load packages
library(tidyverse)
library(modelsummary)
library(tinytable) # simple tables in html, LaTeX, markdown, word, pdf, typst
library(litedown)

# set some session options for `tinytable`
options(tinytable_tt_digits = 3)
# set option to render tinytable math in html
options(tinytable_html_mathjax = TRUE)

# options(tinytable_quarto_disable_processing = FALSE)
# options(tinytable_theme_placement_latex_float = "H")

# load data
load(here::here("data", "research-paper-data-20241029.Rdata"))
# load("data/research-paper-data-20241029.Rdata")

```

```{r}
#| label: fun-logit-gof

# custom function to return a certain selection of logistic regression model
# goodness-of-fit statistics
lrm.gof.stats <- function(model, ...){

  out <- data.frame(
    nobs = stats::nobs(model),
    null.deviance = model$null.deviance,
    df.null = model$df.null,
    logLik = as.numeric(stats::logLik(model)),
    deviance = stats::deviance(model),
    df.residual = stats::df.residual(model),
    chisq = as.numeric(model$null.deviance - stats::deviance(model)),
    df = as.numeric(model$df.null - stats::df.residual(model)),
    'P(>chi)' = pchisq(q=model$null.deviance - stats::deviance(model),
                       df = model$df.null - stats::df.residual(model), 
                       lower.tail = F)
  )
  out <- dplyr::as_tibble(out) |>  
  # rename column variables
  dplyr::rename(
    "Num.Obs."      = "nobs",
    "Log.Lik"       = "logLik",
    "Deviance"      = "deviance",
    "Deviance Null" = "null.deviance",
    "DF"            = "df",
    "chisq"          = "chisq",
    "P(>chisq)"     = "P..chi."
  )
  return(out)
}

```

```{r}

# see here: https://vincentarelbundock.github.io/tinytable/vignettes/theme.html#user-written-themes

theme_mitex <- function(x, ...) {
    fn <- function(table) {
        if (isTRUE(table@output == "typst")) {
          table@table_string <- gsub(
            "\\$(.*?)\\$",
            "#mitex(`\\1`)",
            table@table_string)
        }
        return(table)
    }
    x <- style_tt(x, finalize = fn)
    return(x)
}

```

```{=typst}
#show figure: set block(breakable: true)
```

```{=typst}
#set block(spacing: 2em)
#set par(first-line-indent: (
  amount: 1.5em,
  all: true),
  leading: 0.85em,
  spacing: 1.5em
  )
```

```{r}
#| label: ordinal-logit
#| include: false

# needed to duplicate `group` variable with different column name due to
# conflict with marginaleffects R package
data <- data |>
  mutate(condition = group)

# running ordinal logisitc regression models (proportional odds logistic regression)
m.q25 <- MASS::polr(q25 ~ condition, data = data, method = "logistic", Hess = T)
m.q26 <- MASS::polr(q26 ~ condition, data = data, method = "logistic", Hess = T)
m.q36 <- MASS::polr(q36 ~ condition, data = data, method = "logistic", Hess = T)
m.q37 <- MASS::polr(q37 ~ condition, data = data, method = "logistic", Hess = T)

# "Outcome predicted by a fitted model on a specified scale for a given
# combination of values of the predictor variables, such as their observed
# values, their means, or factor levels (a.k.a. "reference grid")".

# predictions(): unit-level (conditional) estimates.

# predicted probability estimates by treatment or control
q25_preds <- marginaleffects::predictions(m.q25, by = "condition", type = "probs")
q26_preds <- marginaleffects::predictions(m.q26, by = "condition", type = "probs")
q36_preds <- marginaleffects::predictions(m.q36, by = "condition", type = "probs")
q37_preds <- marginaleffects::predictions(m.q37, by = "condition", type = "probs")

# this provides the difference in predicted probabilities between treatment and
# control
pdiff_q25 <- marginaleffects::predictions(
  m.q25,
  by = "condition",
  type = "probs",
  hypothesis = ~ reference | group
)

pdiff_q26 <- marginaleffects::predictions(
  m.q26,
  by = "condition",
  type = "probs",
  hypothesis = ~ reference | group
)

```

```{r}
#| label: tbl-safety-logit
#| tbl-cap: "Ordinal Logistic Regression of Concerns for Violence and Confidence in Voter Safety Regressed on Experiment Condition"
#| tbl-cap-location: top

# save models to a list
safety_models <- list(
  "Concerns for Violence (OR)" = m.q25,
  "Confidence in Voter Safety (OR)" = m.q26)

cm <- c("conditionTreatment" = "Treatment",
        "Very concerned|Somewhat concerned" = "Very concerned | Somewhat concerned",
        "Somewhat concerned|Not too concerned" = "Somewhat concerned | Not too concerned",
        "Not too concerned|Not at all concerned" = "Not too concerned | Not at all concerned",
        "Not at all confident|Not too confident" = "Not at all confident | Not too confident",
        "Not too confident|Somewhat confident" = "Not too confident | Somewhat confident",
        "Somewhat confident|Very confident" = "Somewhat confident | Very confident"
        )

tbl_safety_logit <- modelsummary::modelsummary(
  safety_models,  
  fmt = 3,
  output = "tinytable",
  exponentiate = T, # show odds ratio 
  # coef_omit = c(-4), # only the beta coef, not the intercepts
  # coef_rename = c("conditionTreatment" = "Treatment"),
  coef_map = cm,
  stars = F,
  estimate = c("{estimate}, [{conf.low}, {conf.high}]"),
  statistic = NULL, 
  gof_map = c("nobs", "rmse"),
  note = "Parameter estimate (and confidence intervals) exponentiated to show odds ratios.",
  width = c(2, 1, 1)
  )

tbl_safety_logit |> 
  tinytable::style_tt(
    fontsize = 0.65 # fontsize of text in table in em units (0.75*default)
    ) |> 
  tinytable::style_tt(i = "notes", fontsize = 0.65, align = "l") |> 
  tinytable::theme_tt(theme = theme_mitex)

```

```{r}
#| label: tbl-safety-preds1
#| tbl-cap: "Unit-level predicted probabilities of selecting response or higher comparing experiment condition"

# table that shows the predicted probability of selecting a response or lower
# conditional to the Treatment experiment condition.
tab_safety_preds1 <- q25_preds |>  
  dplyr::as_tibble() |> 
  dplyr::mutate('Pr(>|z|)' = dplyr::case_when(
    p.value < 0.05 ~ '\\<0.05',
    p.value < 0.01 ~ '\\<0.01',
    .default = as.character(p.value)
  )) |> 
  select(group, condition, 
         estimate, std.error, statistic, 'Pr(>|z|)',
         conf.low, conf.high) |> 
  dplyr::rename(
    Condition = condition,
    Violence = group,
    z = statistic) |>
  insight::format_table(digits = 3)

tinytable::tt(
  tab_safety_preds1, digits = 3,
  notes = attr(data$q25, "label")) |> 
  tinytable::style_tt(
    fontsize = 0.65, # fontsize of text in table in em units (0.75*default)
    align = "llccccc") |> 
  tinytable::style_tt(i = "notes", fontsize = 0.65, align = "l")

```

```{r}
#| label: tbl-safety-preds2
#| tbl-cap: "Unit-level predicted probabilities of selecting response or higher conditional to the Treatment experiment condition"

# table that shows the predicted probability of selecting a response or lower
# conditional to the Treatment experiment condition.
tab_safety_preds2 <- q26_preds |> 
  dplyr::as_tibble() |> 
  dplyr::mutate('Pr(>|z|)' = dplyr::case_when(
    p.value < 0.05 ~ '\\<0.05',
    p.value < 0.01 ~ '\\<0.01',
    .default = as.character(p.value)
  )) |> 
  select(group, condition, 
         estimate, std.error, statistic, 'Pr(>|z|)',
         conf.low, conf.high) |> 
  dplyr::rename(
    Condition = condition,
    'Voter Safety' = group,
    z = statistic) |>
  insight::format_table(digits = 3)

tinytable::tt(
  tab_safety_preds2, digits = 3,
  notes = attr(data$q26, "label")) |>  
  tinytable::style_tt(
    fontsize = 0.65, # fontsize of text in table in em units (0.75*default)
    align = "llccccc"
    ) |> 
  tinytable::style_tt(i = "notes", fontsize = 0.65)

```